
```{r}

system.time(
  taxaInLocales <- foreach(i=1:nrow(uTaxaNativity))%dopar% {
    inLocales <- NULL
    for (j in locales){
      if(max(grepl(j, uTaxaNativity$Native.Status[i]))){
        inLocales <- c(inLocales, j)
      }
    }
    return(inLocales)
  })

names(taxaInLocales) <- uTaxaNativity$acceptedTaxonID
## this list is = to taxaInStates #
## No examples for AK #


#generates columns of Presence/Absence


localesList<-list('l48NativeStatusCode'=c('L48'),
                  'akNativeStatusCode'=c('AK'),
                  'hiNativeStatusCode'=c('HI'),
                  'prNativeStatusCode'=c('PR')
                  )

floristicArea<-list('L48', 'AK', 'HI', 'PR')

  
system.time(taxaInLocales2 <- foreach(i=1:nrow(uTaxaNativity))%dopar%{
  floristicArea <- taxaInLocales[[i]]
  inLocales <- NULL
  for (j in locales){
    if (length(which(floristicArea%in%localesList[[j]])) > 0){
      inLocales <- c(inLocales, 'P')
    } else {inLocales <- c(inLocales, 'A')}
  }
  return(inLocales)
})

# floristArea and locales are lists of exactly the same thing, the arguments were circular if not
## written this way

names(taxaInLocales2) <- uTaxaNativity$acceptedTaxonID

#produces a list of P/A (only 'A') with columns (un-named) for each locale

# are these next 2 lines of code necessary? I(k8) commented them out for now
# z <- taxaInDomains
# taxaInDomains <- z
taxaInLocales2 <- data.frame(matrix(unlist(taxaInLocales2), ncol=4, byrow=TRUE))  
names(taxaInLocales2) <- locales
taxaInLocales2$acceptedTaxonID <- uTaxaNativity$acceptedTaxonID

taxaInLocales2 <- data.frame(lapply(taxaInLocales2, as.character), 
                            stringsAsFactors=FALSE)

system.time(
  for (i in 1:nrow(uTaxaNativity)){
    for (a in 1:nrow(taxaInLocales2)){
      #work with records in taxaInLocales2 for relevant taxon:
      if (taxaInLocales2$acceptedTaxonID[a] == uTaxaNativity$acceptedTaxonID[i]){
        #extract native status code for each region (locale)
        for (b in locales) {
          if(max(grepl(b, uTaxaNativity$Native.Status[i]))){
            k <- paste(b, '\\(.*?\\)', sep = ' ')
            l <- gsub("[\\(\\)]", "", 
                      regmatches(uTaxaNativity$Native.Status[i],
                                 regexpr(k, uTaxaNativity$Native.Status[i])))
            print (nativeStatus <- strsplit(l, " ")[[1]][2])
            #replace 'P's in relevant region with correct native status code
            for (j in 1:length(locales)){
              if (length(which(b%in%localesList[[j]])) > 0){
                if (taxaInLocales2[a,j] == 'P'){
                  taxaInLocales2[a,j] <- nativeStatus
                }
              }
            }
          }
        }
      }
    }
  }
)

#KJ: I think this will work once I figure out why P/A = A for everything

usda <- merge(usda, taxaInLocales2, by = "acceptedTaxonID")


## rename field
````

#next repeat for Native/Introduced/ETC
#then leave absences as A, but overwrite presences with nativeStatus

replace Presence data with Native Status Codes

## original code
system.time(
  for (i in 1:nrow(uTaxaNativity)){
    for (a in 1:nrow(taxaInDomains)){
      #work with records in taxaInDomains for relevant taxon:
      if (taxaInDomains$acceptedTaxonID[a] == uTaxaNativity$acceptedTaxonID[i]){
        #extract native status code for each region (locale)
        for (b in locales) {
          if(max(grepl(b, uTaxaNativity$Native.Status[i]))){
            k <- paste(b, '\\(.*?\\)', sep = ' ')
            l <- gsub("[\\(\\)]", "", 
                      regmatches(uTaxaNativity$Native.Status[i],
                                 regexpr(k, uTaxaNativity$Native.Status[i])))
            print (nativeStatus <- strsplit(l, " ")[[1]][2])
            #replace 'P's in relevant region with correct native status code
            for (j in 1:length(domainIDs)){
              if (length(which(b%in%localesInDomains[[j]])) > 0){
                if (taxaInDomains[a,j] == 'P'){
                  taxaInDomains[a,j] <- nativeStatus
                }
              }
            }
          }
        }
      }
    }
  }
)
#end of original Code

```{r replace Presence data with Native Status Codes}

```
