---
title: ""
author: "Katie Jones"
date: \today
urlcolor: blue
output: pdf_document
header includes:
 - \usepackage{longtable}
 - \usepackage{hyperref}
 - \usepackage{ltablex}
---

```{r loadLibraries, echo=FALSE, message=FALSE, warning=FALSE}

library(restR)
library(tidyverse)
library(png)
library(knitr)
library(pander)
library(kableExtra)
library(magrittr)
library(scales)
library(xtable)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")


```


```{r filePaths, echo=FALSE, message=FALSE, warning=FALSE}

if (file.exists(
  'C:/Users/kjones')){
  myPathToNEONplants <- "C:/Users/kjones/Documents/GitHub/neonPlantSampling/stemMapSimulation"
  myPathToFulcrumDev <- "C:/Users/kjones/Documents/GitHub/fulcrumDev/admin_tools_scripts/magpie_requests" 
  myPathToSharepoint <- "C:/Users/kjones/National Ecological Observatory Network, Inc/FSU Only - Plant Docs/protocols/vst_vegetationStructure/dendro_prototype"
}

```


```{r loadData, echo=FALSE, message=FALSE, warning=FALSE}

df <- read.csv(paste(myPathToSharepoint, "vst_data_d.csv", sep="/"), stringsAsFactors=F )
## BLAN KONZ data
#df <- read.csv(paste(myPathToSharepoint, "vst_KONZ_BLAN_tower.csv", sep="/"), stringsAsFactors = F)

nlcd_optimization <- read.csv(paste(myPathToSharepoint, "optimizedPlots_sub.csv", sep="/"), 
                              stringsAsFactors=F)


equations <- read.csv(paste(myPathToSharepoint, 'Final_list/chojnacky_assignment.csv', sep="/"), 
                      stringsAsFactors = F)
parameters <- read.csv(paste(myPathToSharepoint, 'Final_list/chojnacky_parameters.csv', sep="/"), 
                       stringsAsFactors = F)


```


```{r subset dataframe, echo=FALSE, message=FALSE, warning=FALSE}

df$plotEvent <- paste(df$plotid, df$eventid, sep="_")

#opti_plot <- unique(nlcd_optimization$plotID)

lowMorton <- NULL
maxEvent <- NULL
sites <- unique(df$siteid)
#sites <- 'BONA'
#sites <- c('BONA', 'NIWO', 'DEJU')
#sites <- c('BLAN', 'KONZ')
plots <- sort(unique(df$plotid))

for (i in 1:length(plots)){
  sub <- filter(df, plotid==plots[i])
  maxEvent <- c(maxEvent, max(sort(sub$plotEvent)))
  rm(sub)
}


df_sub <- df%>%
  filter(siteid%in%sites & plantstatus%in%c("1", "4", "5", "6", "7", "9", 
                                           "Live", "Live, disease damaged", "Live, insect damaged", "Live, other damage", "Live, physically damaged" ), plotEvent%in%maxEvent)%>% 
  arrange(plotid, desc(eventid))

### suspect stem diameter at BONA, reset
df_sub$stemdiameter[df_sub$siteid=='BONA'& df_sub$stemdiameter==120] <- 12.0
df_sub <- df_sub[!duplicated(df_sub$individualid), ]


for (i in 1:length(sites)){
   sub <- filter(df, siteid==sites[i]& plantstatus%in%c("1", "4", "5", "6", "7", "9", 
                                           "Live", "Live, disease damaged", "Live, insect damaged", "Live, other damage", "Live, physically damaged" ),)
   lowMorton <- c(lowMorton, sort(unique(sub$plotid))[1:5])
   rm(sub)
 }

```

#### New script

```{r calculate biomass, echo=FALSE}


#ln(biomass_kg)= b0 + b1 * ln(diameter_cm)

eq <- select(equations, allometry_ID, taxonid=taxonID)
pm <- select(parameters, allometry_ID, b0, b1, minDiameter, maxDiameter)


#df <- left_join(df, eq)

df_sub$agb <- NA

#i=49

for(i in 1:nrow(df_sub)){
  ch_allID <- ifelse(is.na(df_sub$taxonid[i])|df_sub$taxonid[i]=='', "H7", eq$allometry_ID[eq$taxonid==df_sub$taxonid[i]])
  temp.b0 <- pm$b0[pm$allometry_ID==ch_allID]
  temp.b1 <- pm$b1[pm$allometry_ID==ch_allID]
  df_sub$agb[i] <- round(exp(temp.b0 + temp.b1*log(df_sub$stemdiameter[i])), digits=1)
}


## try mutate or apply
## make function first


```


# Plot and Stem Selection

The 5 lowest morton order plots are targeted for annual vegetation structure measurements

## Assumptions for tree selection for dendrometer band installation:

* all stemDiameters converted to aboveground biomass based on specific allometric equations from Chojnacky et al., 2014
* within subset of 5 target productivity Tower plots all trees > 1 cm & < 75th percentile aboveground biomass located in subset of plots measured with diameter tapes or calipers on an annual basis (Table 2)
* all trees > 75th  percentile aboveground biomass located in subset of plots recieve dendrometer bands 
* all trees $\geq$ 95th percentile aboveground biomass in all Tower plots recieve dendrometer bands

The diameter thresholds, based on aboveground biomass are summarized in Table 1

```{r determine thresholds, echo=FALSE}

plots <- data.frame("plot"= sort(unique(df_sub$plotid)), stringsAsFactors = FALSE)


thresh <- df_sub%>%
  group_by(siteid)%>%
  summarize(t2 = round(quantile(stemdiameter, 0.75), 1),
            t3 = round(quantile(stemdiameter, 0.95), 1),
            b2 = quantile(agb, 0.75),
            b3 = quantile(agb, 0.95),
            b2_d = min(stemdiameter[agb>=b2]),
            b3_d = min(stemdiameter[agb>=b3])
)


```


```{r comaparisonTable, echo=FALSE, message=FALSE, warning=FALSE}

dbh_agb_thresh <- thresh%>%
  select(siteid, "75th"=b2_d, "95th"=b3_d)#%>%

```

```{r compare dendro plot size distribution to opti plot size distribution, echo=FALSE, }

   low <- df_sub[df_sub$plotid%in%lowMorton,]
  opti <- df_sub[df_sub$plotid%in%opti_plot,]


 out_plot <- ggplot(data=df_sub, aes(x=stemdiameter))+
  facet_wrap(~siteid)+
  geom_density(color="black", size=1)+
  geom_density(data=df_sub[df_sub$plotid %in% lowMorton,], color="orange", size=1) +
  geom_density(data=df_sub[df_sub$plotid %in% opti_plot,], color= "blue", size=1) +
  #xlim(10, 40)+
  ylim(0, 0.15)+
  #geom_vline(data=dbh_agb_thresh, aes(xintercept=as.numeric(`75th`)),
  #           color="black", linetype="dashed", size=0.5)+
  ggtitle("Stem diameter comparison: Black = all plots, Blue = Dendrometer plots, Orange = Optimized plots") #+
  # geom_label(aes(x = 20, y = 0.1, 
  #                  label = paste("n=", nrow(df_sub), "\n",
  #                                                "dendro stems =", nrow(df_sub[plotid%in%lowMorton,]), "\n",
  #                                                "optimize stem =", nrow(df_sub[plotid%in%opti_plot,]),
  #                                                sep=""),
  #                   hjust="left"),
  #            group=df_sub$siteid)




```

```{r add columns for materials}
df_sub$addDBH <- ifelse(df_sub$stemdiameter<50, 4, 7.5)

df_sub$bandLength <- ceiling((df_sub$stemdiameter+df_sub$addDBH)*3.14)

df_sub$springSize <- ifelse(df_sub$stemdiameter < 50, "76mm", "127mm")

```




```{r summary table}

df_sumTable <- df_sub%>%
  group_by(siteid)%>%
  summarise(n_allPlots=n(),
            n_d_prod=sum(low),
            n_optimized=sum(op))


df_d_bands <- NULL
i='BONA'
for (i in unique(df_sub$siteid)){
  siteid <- i
  thresh1 <- dbh_agb_thresh$`75th`[dbh_agb_thresh$siteid==i]
  thresh2 <- dbh_agb_thresh$`95th`[dbh_agb_thresh$siteid==i]
  df_orig <- df_sub[df_sub$siteid==i & df_sub$plotid%in%lowMorton &
                     df_sub$stemdiameter>thresh1 | df_sub$siteid==i &
                     !df_sub$plotid%in%lowMorton & df_sub$stemdiameter>thresh2,]
  df_opti <- df_sub[df_sub$siteid==i & df_sub$plotid%in%opti_plot &
                      df_sub$stemdiameter>thresh1 | df_sub$siteid==i &
                      !df_sub$plotid%in%opti_plot & df_sub$stemdiameter>thresh2,]
  bands_orig <- nrow(df_orig[df_orig$siteid==i,])
  plots_orig <- length(unique(df_orig$plotid[df_orig$siteid==i]))
  bands_optimized <- nrow(df_opti[df_opti$siteid==i,])
  plots_optimized <- length(unique(df_opti$plotid[df_opti$siteid==i]))
  df_d_bands = rbind(df_d_bands, data.frame(siteid, bands_orig, plots_orig, bands_optimized, plots_optimized))
}  
  
df_sumTable <- left_join(df_sumTable, df_d_bands)

# kable(df_sumTable, "html") %>%
#   kable_styling(bootstrap_options = c("striped", "hover"), full_width = F) %>%
#   save_kable("countSummary_blanKonz.html")


```
# Pull vst data for BLAN & KONz


```{r get materials cost}

tapeCost <- 8.31
springCost <- 9.55  #small size springs

materials_orig <- df_orig%>%
  group_by(siteid)%>%
  summarise(n=n(), totalBandLength=round(sum(bandLength),digits = 0),
            tapeCount =round(ceiling(totalBandLength/640), digits=0),
            springPacks = ceiling(n/10))

materials_orig$cost <- (materials_orig$tapeCount*1.5*tapeCost)+(materials_orig$springPacks*1.5*springCost)

# $40/hr * 10 min/band
materials_orig$labor <- materials_orig$n*(40*0.18)

materials_orig$total <- materials_orig$cost + materials_orig$labor

materials_opti <- df_opti%>%
  group_by(siteid)%>%
  summarise(n=n(), totalBandLength=round(sum(bandLength),digits = 0),
            tapeCount =round(ceiling(totalBandLength/640), digits=0),
            springPacks = ceiling(n/10))

materials_opti$cost <- (materials_opti$tapeCount*1.5*tapeCost)+(materials_opti$springPacks*1.5*springCost)

materials_opti$labor <- materials_opti$n*(40*0.18)

materials_opti$total <- materials_opti$cost + materials_opti$labor

costDiff_bona <- materials_opti$total[materials_opti$siteid=='BONA']-materials_orig$total[materials_orig$siteid=='BONA']


```

