---
title: "Dendrometer Cost Estimate, 2019 & 2020"
author: "Katie Jones"
date: /today
urlcolor: blue
output: pdf_document
header includes:
 - \usepackage{longtable}
 - \usepackage{hyperref}
 - \usepackage{ltablex}
 - \usepackage{rotating}
 - \usepackage{pdflscape}
---
```{r loadLibraries, echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(png)
library(knitr)
library(pander)
library(kableExtra)
library(magrittr)
library(scales)
library(xtable)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")


```



```{r filePaths, echo=FALSE, message=FALSE, warning=FALSE}

if (file.exists(
  'C:/Users/kjones')){
  myPathToNEONplants <- "C:/Users/kjones/Documents/GitHub/neonPlantSampling/stemMapSimulation"
  myPathToFulcrumDev <- "C:/Users/kjones/Documents/GitHub/fulcrumDev/admin_tools_scripts/magpie_requests" 
  myPathToSharepoint <- "C:/Users/kjones/National Ecological Observatory Network, Inc/FSU Only - Plant Docs/protocols/vst_vegetationStructure/dendro_prototype"
}


```




```{r loadData, echo=FALSE, message=FALSE, warning=FALSE}

df <- read.csv(paste(myPathToSharepoint, "vst_data_d.csv", sep="/"), stringsAsFactors=F )
equations <- read.csv(paste(myPathToSharepoint, 'Final_list/chojnacky_assignment.csv', sep="/"), stringsAsFactors = F)
parameters <- read.csv(paste(myPathToSharepoint, 'Final_list/chojnacky_parameters.csv', sep="/"), stringsAsFactors = F)
map <- read.csv(paste(myPathToSharepoint, 'nineSite_map.csv', sep='/'), stringsAsFactors = F)
labor <- read.csv(paste(myPathToSharepoint, "laborEstimates.csv", sep="/"),stringsAsFactors=F )
labor$siteid <- trimws(labor$siteid)

```



```{r subsetData, echo=FALSE, message=FALSE, warning=FALSE}

df$plotEvent <- paste(df$plotid, df$eventid, sep="_")
df$indEvent <- paste(df$individualid, df$eventid, sep="_")

maxEvent <- NA
sites <- unique(df$siteid)
plots <- sort(unique(df$plotid))

for (i in 1:length(plots)){
  sub <- filter(df, plotid==plots[i])
  maxEvent <- c(maxEvent, max(sort(sub$plotEvent)))
  rm(sub)
}

#site <- "WREF"
#site <- c("RMNP")
site <- c("NIWO", "RMNP", "SJER", "BONA", "HEAL", "DEJU", "YELL", "SOAP", "TEAK", "WREF")


df_sub <- df%>%
  filter(siteid%in%site & plantstatus%in%c("1", "4", "5", "6", "7", "9", 
                                           "Live", "Live, disease damaged", "Live, insect damaged", "Live, other damage", "Live, physically damaged" ), plotEvent%in%maxEvent)%>% 
  arrange(plotid)



```


```{r calculate biomass, echo=FALSE}


#ln(biomass_kg)= b0 + b1 * ln(diameter_cm)

eq <- select(equations, allometry_ID, taxonid=taxonID)
pm <- select(parameters, allometry_ID, b0, b1, minDiameter, maxDiameter)

#df <- left_join(df, eq)

df_sub$agb <- NA

for(i in 1:nrow(df_sub)){
  ch_allID <- ifelse(is.na(df_sub$taxonid[i])|df_sub$taxonid[i]=='', "H7", eq$allometry_ID[eq$taxonid==df_sub$taxonid[i]])
  temp.b0 <- pm$b0[pm$allometry_ID==ch_allID]
  temp.b1 <- pm$b1[pm$allometry_ID==ch_allID]
  df_sub$agb[i] <- round(exp(temp.b0 + temp.b1*log(df_sub$stemdiameter[i])), digits=1)
}


## try mutate or apply
## make function first


```


```{r distribution based on biomass vs stemdiameter, echo=FALSE}

##stemdiameter
# ggplot(data=df_sub, aes(x=stemdiameter))+
#   geom_density(fill="gray")+
#   geom_vline(data=df_sub, aes(xintercept=quantile(stemdiameter, .75)),
#              color="blue", linetype="dashed", size=1) +
#   geom_vline(data=df_sub, aes(xintercept=quantile(stemdiameter, .95)),
#              color="orange", linetype="dashed", size=1) +
#   ggtitle("Distribution of dbh")+
#   geom_label(aes(x = 40, y = 0.02, label = paste("n= ", nrow(df_sub), "\n",
#                                                  "n >=75th percentile= ", nrow(df_sub[df_sub$stemdiameter >= quantile(stemdiameter, .75),]), "\n",
#                                                  "n >95th percentile= ", nrow(df_sub[df_sub$stemdiameter >= quantile(stemdiameter, .95),]))))

# 
# 
# ggplot(data=df_sub, aes(x=agb))+
#   geom_density(fill="gray")+
#   geom_vline(data=df_sub, aes(xintercept=quantile(agb, .75)),
#              color="blue", linetype="dashed", size=1) +
#   geom_vline(data=df_sub, aes(xintercept=quantile(agb, .95)),
#              color="orange", linetype="dashed", size=1) +
#   ggtitle("Distribution of above ground biomass") +
#   geom_label(aes(x = 5000, y = 0.0005, label = paste("n= ", nrow(df_sub), "\n",
#                                                  "n >=75th percentile= ", nrow(df_sub[df_sub$agb>= quantile(agb, .75),]), "\n",
#                                                  "n >95th percentile= ", nrow(df_sub[df_sub$agb >= quantile(agb, .95),]))))+
#   ylim(0, 0.001)



```

# Plot Selection

## Assumptions for tree selection for dendrometer band installation:

* all stemDiameters converted to aboveground biomass based on specific allometric equations from Chojnacky et al., 2014
* within subset of 5 target Tower plots all trees > 1 cm & < 75th percentile aboveground biomass located in subset of plots measured with diameter tapes or calipers on an annual basis (Table 2)
* all trees > 75th  percentile aboveground biomass located in subset of plots recieve dendrometer bands 
* all trees $\geq$ 95th percentile aboveground biomass in all Tower plots recieve dendrometer bands

The diameter thresholds at each site, based on aboveground biomass are summarized in Table 1

```{r determine thresholds, echo=FALSE}

plots <- data.frame("plot"= sort(unique(df_sub$plotid)), stringsAsFactors = FALSE)


thresh <- df_sub%>%
  group_by(siteid)%>%
  summarize(t2 = round(quantile(stemdiameter, 0.75), 1),
            t3 = round(quantile(stemdiameter, 0.95), 1),
            b2 = quantile(agb, 0.75),
            b3 = quantile(agb, 0.95),
            b2_d = min(stemdiameter[agb>=b2]),
            b3_d = min(stemdiameter[agb>=b3])
)


```


```{r comaparisonTable, echo=FALSE, message=FALSE, warning=FALSE}

dbh_agb_thresh <- thresh%>%
  select(siteid, "75th"=b2_d, "95th"=b3_d)#%>%
#  filter(siteid!="WREF")

panderOptions('table.split.table', Inf)
 pander::pander(dbh_agb_thresh, justify="left",  missing='',  caption="stemDiameter thresholds for the 75th and 95th percentiles based on aboveground biomass")
 

#write.csv(dbh_agb_thresh, paste("bandlists/", "bandDiameter_summary_2019.csv", sep=""), row.names=FALSE)

#
# dbh_agb <- matrix(siteid,c(thrt2, b2_d, t3, b3_d), ncol=2, byrow=TRUE)
# colnames(dbh_agb) <- c("diameter", "biomass")
# rownames(dbh_agb) <- c("75th","95th")
# 
# dbh_agb <- as.table(dbh_agb)
# dbh_agb

```


```{r Band deets, echo=FALSE, anchor="Table", message=FALSE}


df_sub$addDBH <- ifelse(df_sub$stemdiameter<10, 2.5,
                        ifelse(df_sub$stemdiameter < 50, 4, 7.5))

for (i in 1:length(unique(df_sub$siteid))){

  df_sub$stemRank <- rank(df_sub$agb, ties.method="first")
}


for (i in 1: nrow(unique(df_sub))){
  df_sub$banded_sub[i] <- ifelse(df_sub$agb[i]>=thresh$b2[thresh$siteid==df_sub$siteid[i]], "Y", "N")
  df_sub$banded_all[i] <- ifelse(df_sub$agb[i]>=thresh$b3[thresh$siteid==df_sub$siteid[i]], "Y", "N")
}

################## fix this to reset every other stem rank to N ######################

# 
# even <- df_sub$stemRank[which(df_sub$stemRank %% 2 != 0)]
# df_sub$banded_sub[df_sub$siteid=="RMNP" & df_sub$banded_sub=="Y" & df_sub$banded_all=="N" & df_sub$stemRank%in%even] <- "N"
# 


df_sub$springLength <- ifelse(df_sub$banded_sub=="Y", 
                              ifelse(df_sub$stemdiameter < 10, "short", 
                                     ifelse(df_sub$stemdiameter > 10 & df_sub$stemdiameter < 50, "long_0.26", "long_0.31")),
                              NA)
df_sub$templateSize <- ifelse(df_sub$banded_sub=="Y", 
                              ifelse(df_sub$stemdiameter < 10, "sm", 
                                     ifelse(df_sub$stemdiameter > 10 & df_sub$stemdiameter < 50, "md", "lg")),
                              NA)



```


```{r target plot selection, echo=FALSE, anchor="Table", message=FALSE}

lowMorton <- NA
maxEvent <- NA
sites <- unique(df_sub$siteid)
plots <- unique(df_sub$plotid)
key <- NA


pl <- data.frame(x = 1:5)

for (i in 1:length(unique(df_sub$siteid))){
  sub <- filter(df_sub, siteid==sites[i] & banded_sub=="Y")
  p <-   if(sub$siteid[i]=="WREF"){
    sort(unique(sub$plotid))[4:8]
  }else{
    sort(unique(sub$plotid))[1:5]
  }
  t <- as.data.frame(p)
  colnames(t) <- sub$siteid[1]
  pl <- bind_cols(pl, t)
  lowMorton <- c(lowMorton, p)
  maxEvent <- c(maxEvent, max(sort(sub$eventid)))
  rm(sub)
}

# pl <- pl%>%
#   select(-x, -WREF)

panderOptions('table.split.table', Inf)
pander::pander(pl, justify="left",  missing='', caption="Target plots by site")

#write.csv(pl,  "vstFullPlots.csv", row.names=FALSE)

```




```{r make final list, echo=FALSE, message=FALSE}

# assign domains and  group by DSF
df_sub$domainid <- substr(df_sub$individualid, 10, 12)
df_sub$domainid <- ifelse(df_sub$domainid%in%c("D10", "D13"), "D10/13", 
                          ifelse(df_sub$domainid%in%c("D18", "D19"), "D18/19", df_sub$domainid))


# subset to selected trees 
df_out <- df_sub%>%
  filter(plotid%in%lowMorton & banded_sub=="Y" | banded_all=="Y")%>%
  left_join(map)%>%
  select(domainid, siteid, plotid, subplotid, pointid, stemazimuth, stemdistance, taxonid, individualid, growthform, stemdiameter, templateSize, springLength)%>%
  arrange(domainid, plotid, subplotid, desc(stemdiameter))

# write files for each site

 for (i in 1:length(sort(unique(df_out$siteid)))){
   sub <- df_out[df_out$siteid==sort(unique(df_out$siteid))[i],]
   #write.csv(sub, paste("bandlists/", unique(sub$siteid), "_treesToBand.csv", sep=""), row.names=FALSE)
   rm(sub)
 }


```

# Equipment

## Assumptions for equipment lists:

* totalBandLenth = (dbh + extra, based on size) * 3.14
* label tape = sum tapeCount (totalBandLenght/640)*1.2
* springs 1.5 in = only used for stems < 10 cm dbh, no qualifying stems (may be used during remeasurement to extend already installed bands)
* springs 3 in 0.26 guage = count of stems 10-40 cm dbh *1.25 
* springs 3 in 0.31 guage = count of stems >50 cm dbh *1.25 
* hole punch, tin snips, hatchet = 1 per DSF
* mora knife = 2 per DSF (may need additional knives if multiple sites within a domain are installing concurrently - D18/19)
* metal digital calipers - model listed here suggested in Smithsonian protocols, comparable product is acceptable. May need to use dial calipers intead of digital in wet environments (mcmaster.com item #	2325A55)

* TOTAL_known = sum of listed equipment costs + 2.5% fully burdened cost
* TOTAL_est = incresed equipment cost by 40% to account for sites with no initial data

\textbf{Additional Consideration:} _per unit cost of springs varies from \$1.50- $7, depending on volume, if ordered directly by domains. This estimate assumes ordering by HQ then distributing to DSFs. It may still be preferable, logistically, to order direct from supplier, despite increased cost._ 

```{r materials, echo=FALSE, message=FALSE, warning=FALSE}

options(knitr.kable.NA = '')


materials <- df_sub%>%
  filter(banded_sub=="Y"&plotid%in%lowMorton |banded_all =="Y" )%>%
  filter(siteid!="WREF")%>%
  group_by(domainid, siteid, springLength)%>%
  summarise(n=n(), maxDBH=max(stemdiameter), totalBandLength=ceiling(sum((stemdiameter+addDBH)*3.14)),
            tapeCount =round(ceiling(totalBandLength/640)*1.2, digits=0),
            springs = round(ceiling(n*1.25), digits=0))

panderOptions('table.split.table', Inf)
 pander::pander(materials, justify="left",  missing='', split.cells = c(12, 50, 5), caption="Banding summary by site")

 materials_sub <- select(materials, -maxDBH, -totalBandLength, -n)
 
# kable(materials_sub, "html", escape = FALSE) %>%
#    kable_styling(bootstrap_options = c("striped", "hover"), full_width = F) %>%
#    save_kable("2019_materials_list.html")


buy <- data.frame("Vendor"=c(text_spec("www.mcmaster.com", link = "https://bit.ly/2ttVfDW"),
                         text_spec("www.leespring.com", link = "https://www.leespring.com/extension-springs?search=LE026C05S"),
                        text_spec("www.leespring.com", link = "https://www.leespring.com/extension-springs?search=LE026C11S"),
                        text_spec("www.leespring.com", link = "https://www.leespring.com/extension-springs?search=LE031C11S"),
                        text_spec("www.roperwhitney.com", link = "https://bit.ly/2MtX4gF"),
                        text_spec("www.amazon.com", link = "https://amzn.to/2NgBmcy"),
                        text_spec("www.knifecenter.com", link = "https://bit.ly/2XolfNE"),
                        text_spec("www.forestry-suppliers.com", link = "https://bit.ly/2TjZvEE}"),
                        text_spec("www.mcmaster.com", link= "https://bit.ly/2NusGTT")),
                  "Item"=as.character(c("item no. 1598T62 - stainless steel label tape 1/2 (1.3 cm) in 21ft (6.4 m) .007 thickness",
                           "lee Spring - LE 026C 05 S - 1.5 in springs, stainless steel",
                           "lee Spring - LE 026C 11 S - 3 in springs, stainless steel 0.26 gauge",
                           "Lee Spring - LE 031C 11 S - 3 in springs, stainless steel 0.31 gauge",
                           "heavy duty hole punch",
                           "tin snips",
                           "mora knife", 
                           "hatchet - use CDW equipment if available",
                           "mcMaster item no. #2325A55")) #8647A42 - metal calipers"))
                  )

buy$count <- ifelse(buy$Item%in%c("heavy duty hole punch", "tin snips"), length(unique(df_sub$domainid)), NA)
buy$count <- ifelse(buy$Item=="item no. 1598T62 - stainless steel label tape 1/2 (1.3 cm) in 21ft (6.4 m) .007 thickness",
                    sum(materials$tapeCount),
                    buy$count)
buy$count <- ifelse(buy$Item=="lee Spring - LE 026C 05 S - 1.5 in springs, stainles steel",
                    materials$springs[materials$springLength=='short'],
                    buy$count)
buy$count <- ifelse(buy$Item=="lee Spring - LE 026C 11 S - 3 in springs, stainless steel 0.26 gauge",
                    sum(materials$springs[materials$springLength=='long_0.26']),
                    buy$count)
buy$count <- ifelse(buy$Item=="Lee Spring - LE 031C 11 S - 3 in springs, stainless steel 0.31 gauge",
                    sum(materials$springs[materials$springLength=='long_0.31']),
                    buy$count)
buy$count <- ifelse(buy$Item=="mora knife", round(2*length(unique(df_sub$domainid)), digits = 0),
                    buy$count)
buy$count<- ifelse(buy$Item=="hatchet - use CDW equipment if available", round(length(unique(df_sub$domainid)), digits=0),
                    buy$count)
buy$count <- ifelse(buy$Item=="mcMaster item no. 	8647A42 - metal calipers", 2*length(unique(df_sub$domainid)), 
                    buy$count)


## price
buy$price <- ifelse(buy$Item=="heavy duty hole punch", buy$count * 98,NA)
buy$price <- ifelse(buy$Item=="tin snips", buy$count*20,  buy$price)
buy$price <- ifelse(buy$Item=="item no. 1598T62 - stainless steel label tape 1/2 (1.3 cm) in 21ft (6.4 m) .007 thickness",
                    buy$count*8.31, buy$price)
buy$price <- ifelse(buy$Item=="lee Spring - LE 026C 05 S - 1.5 in springs, stainles steel",
                    buy$count*1.3,  buy$price)
buy$price <- ifelse(buy$Item=="lee Spring - LE 026C 11 S - 3 in springs, stainless steel 0.26 gauge",
                    buy$count*1.1, buy$price) # $1.10 a piece for 600-700
buy$price <- ifelse(buy$Item=="lee Spring - LE 031C 11 S - 3 in springs, stainless steel 0.31 gauge",
                    buy$count*4.1,  buy$price) #$4 a piece for 20-50
buy$price <- ifelse(buy$Item=="mora knife",
                    buy$count*15,  buy$price)
buy$price <- ifelse(buy$Item=="hatchet - use CDW equipment if available",
                    buy$count*40,  buy$price)
buy$price <- ifelse(buy$Item=="mcMaster item no. 	8647A42 - metal calipers",
                    buy$count*137,  buy$price)

newRow <- data.frame("TOTAL_known", NA, NA, round(sum(buy$price, na.rm= TRUE), digits=2)*1.025)
names(newRow) <- names(buy)
buy <- bind_rows(buy, newRow)

newRow <- data.frame( "TOTAL_est.", "TOTAL_known * 1.2, to account for YELL with no initial data", NA, buy$price[buy$Vendor=="TOTAL_known"]*1.2)
names(newRow) <- names(buy)
buy <- bind_rows(buy, newRow)


buy$price <- ifelse(!is.na(buy$price),dollar(buy$price), NA)

buy_sub <- select(buy, -count, -price)

# kable(buy_sub, "html", escape = FALSE) %>%
#   kable_styling(bootstrap_options = c("striped", "hover"), full_width = F) %>%
#   save_kable("2019_equipment_vendors.html")

panderOptions('table.split.table', Inf)
 pander::pander(buy, justify="left",  missing='', split.cells = c(12, 50, 5), caption="Equipment list for banding stems at sites with initial vst data available")

 
#   split.cells = c(20, 50, 5)
# \newcolumntype{g}{>{\columncolor{Gray}}c}
# \begin{tabular}{lll}
# \textbf{Vendor} & \textbf{Item} & \textbf{Count}
#   \hline
#   `r paste( "\\\\{", buy$Vendor, "} &",
#         buy$Item, " &", buy$count, " \\\\ ","\\hline", sep="", collapse=" ")`
# \label{tab-plotLists}
# \end{tabular}
# \caption{Equipment list}
#  
 
```

# Labor and Travel

## Assumptions for labor cost estimates:

* hrs_per_plot = average time for 2 staff to complete full vst measurement of a Tower plot (provided by DM)
* n_targetPlot = count of plots in the lowest 5 mortorn order list
* n_nonTarget = count of plots that contain largest trees
* n_bands = count of trees >75th percentile of biomass in target plots + count of trees >95th percentile of biomass in non-target plots
* hrs_field_install = (n_targetPlot \* hrs_per_plot  \* 2 staff) + (n_nonTarget plots \* 30 min travel) + (n_bands \* 25 mins(WREF, SJER, and SOAP), 12 mins (all other sites)
* hrs_field_remeasure = (n_targetPlot \* hrs_per_plot  \* 2 staff) + (n_nonTarget plots \* 30 min travel) + (n_bands \* 3 mins (all sites) 
* hrs_travel = (hrs_field / 8)\*(hrs_to_site * 2) _assumed 0.5 hrs travel for all sites with field lodging (i.e. DEJU, HEAL, SOAP, TEAK)_
* labor_cost = (hrs_field + hrs_travel) \* \$51.66/hr (2019 sites) OR \$53/hr (2020 sites: SOAP, TEAK, YELL)
* travel_cost = ((hrs_field / 8) * lodging) + ((hrs_field / 8) * perDiem) * 15% GRA

```{r fs_laborEstimate, message=FALSE, echo=FALSE, warning=FALSE, results="asis"}
options(knitr.kable.NA = '')

labor <- select(labor, siteid, hrs_per_plot, travel_to_site, lodging, perDiem)

labor_out <- df_out%>%
  left_join(labor, by='siteid')%>%
  group_by(siteid, hrs_per_plot, travel_to_site, lodging, perDiem)%>%
  summarise(n_targetPlots=ifelse(length(unique(plotid))<5, length(unique(plotid)), 5),  
            n_nonTargetPlots=ifelse(length(unique(plotid))-5 <1, 0,length(unique(plotid))-5)) #[targetPlot=="N"])))


for(i in 1:length(labor_out$siteid)){
  temp <- materials[materials$siteid==labor_out$siteid[i],]
  labor_out$n_bands[i] <- sum(temp$n)
  rm(temp)
}

labor_out <- ungroup(labor_out)

labor_out <- add_row(labor_out, siteid="YELL*",
                     hrs_per_plot=labor$hrs_per_plot[labor$siteid=="YELL"],
                     travel_to_site=labor$travel_to_site[labor$siteid=="YELL"],
                     lodging=labor$lodging[labor$siteid=="YELL"],
                     perDiem=labor$perDiem[labor$siteid=="YELL"],
                     n_targetPlots=5,
                     n_nonTargetPlots=5, n_bands=100)



labor_out$install_time <- (labor_out$n_targetPlots*(labor_out$hrs_per_plot*2))+(labor_out$n_nonTargetPlots*0.5)+(labor_out$n_bands * ifelse(labor_out$siteid%in%c("WREF", "SJER", "SOAP"), 0.41, 0.2))

labor_out$remeasure_time <- (labor_out$n_targetPlots*(labor_out$hrs_per_plot*2))+(labor_out$n_nonTargetPlots*0.5)+(labor_out$n_bands * 0.05)

labor_out$personDays_install <- ceiling(labor_out$install_time/8)
labor_out$personDays_remeasure <- ceiling(labor_out$remeasure_time/8)

labor_out$travelTime_install <- NA
labor_out$labor_cost_install <- NA
labor_out$travel_cost_install <- NA

labor_out$travelTime_remeasure <- NA
labor_out$labor_cost_remeasure <- NA
labor_out$travel_cost_remeasure <- NA


for(i in 1:length(labor_out$siteid)){
  temp <- labor[labor$siteid==substr(labor_out$siteid[i], 1, 4),]
  labor_out$travelTime_install[i] <- labor_out$personDays_install[i]*(temp$travel_to_site*2)
  labor_out$labor_cost_install[i] <- round((labor_out$install_time[i]+labor_out$travelTime_install[i]) * ifelse(labor_out$siteid[i]%in%c("YELL", "SOAP", "TEAK"), 53, 51.66), digits =2)
  labor_out$travel_cost_install[i] <- ifelse(!is.na(temp$lodging), 
                                     ((temp$lodging*labor_out$personDays_install[i])+(temp$perDiem *labor_out$personDays_install[i]))*1.15, 
                                     0)
  labor_out$travelTime_remeasure[i] <- labor_out$personDays_remeasure[i]*(temp$travel_to_site*2)
  labor_out$labor_cost_remeasure[i] <- round((labor_out$remeasure_time[i]+labor_out$travelTime_remeasure[i]) * 53, digits =2)
  labor_out$travel_cost_remeasure[i] <- ifelse(!is.na(temp$lodging), 
                                     ((temp$lodging*labor_out$personDays_remeasure[i])+(temp$perDiem *labor_out$personDays_remeasure[i]))*1.15, 
                                     0)
  rm(temp)
}

labor_out$totalCost_install <- labor_out$labor_cost_install+labor_out$travel_cost_install
labor_out$totalCost_remeasure <- labor_out$labor_cost_remeasure+labor_out$travel_cost_remeasure


assumptions <- labor_out%>%
  mutate(hrs_per_band_install=ifelse(labor_out$siteid%in%c("WREF", "SJER", "SOAP*"), 0.37, 0.18), hrs_per_band_remeasure= 0.05)%>%
  select(siteid, hrs_plot=hrs_per_plot, n_target=n_targetPlots, n_nonTarget=n_nonTargetPlots, n_bands, hrs_band=hrs_per_band_install, hrs_to_site=travel_to_site, lodging, perDiem)

##summarize installation costs
install_costs <- labor_out%>%
  filter(siteid!="WREF")%>%
  select(siteid, hrs_field=install_time, hrs_travel=travelTime_install, labor_cost=labor_cost_install, travel_cost=travel_cost_install, total_cost_install=totalCost_install)
  
newRow <- data.frame("TOTAL", NA, NA, NA, NA, round(sum(install_costs$total_cost_install), digits=2))
names(newRow) <- names(install_costs)

install_costs <- bind_rows(install_costs, newRow)
install_costs$total_cost <- dollar(install_costs$total_cost_install)

###create the same table for remeasurement
annual_costs <- labor_out%>%
  select(siteid, hrs_field=remeasure_time, hrs_travel=travelTime_remeasure, labor_cost=labor_cost_remeasure, travel_cost=travel_cost_remeasure, total_cost_annual=totalCost_remeasure)
  
newRow <- data.frame("TOTAL", NA, NA, NA, NA, round(sum(annual_costs$total_cost_annual), digits=2))
names(newRow) <- names(annual_costs)

annual_costs <- bind_rows(annual_costs, newRow)
annual_costs$total_cost_annual <- dollar(annual_costs$total_cost_annual)


panderOptions('table.split.table', Inf)
 pander::pander(assumptions, justify="center",  missing='',  caption="Assumptions for cost development. * YELL, baseline data for are not yet available, labor estimates are based on data from similar sites.")

panderOptions('table.split.table', Inf)
 pander::pander(install_costs, justify="center",  missing='',  caption="Labor and travel costs for dendrometer band installation and full vst meaurement of 5 target plots.")
 
 panderOptions('table.split.table', Inf)
 pander::pander(annual_costs, justify="center",  missing='',  caption="Labor and travel costs for full vst measurement of 5 target plots and re-meaurement of all banded trees.")
```

Materials and labor total cost, year 1 = **`r dollar(as.numeric(gsub('[$,]', '', buy$price[buy$Vendor=="TOTAL_est."]))+as.numeric(gsub('[$,]', '', install_costs$total_cost[install_costs$siteid=="TOTAL"])))`**

Annual re-measurement labor total cost for years not currently on vst schedule, this estimate includes Wind River and is based on 2020 labor rates = **`r annual_costs$total_cost_annual[annual_costs$siteid=="TOTAL"]`**

# Timeline

```{r, message=FALSE, echo=FALSE, warning=FALSE, results="asis"}
site <- c(site, "WREF")

timeline <- data.frame("siteid"= site)
                         
timeline$proposedYear <- ifelse(timeline$siteid%in%c("SOAP", "TEAK"), "2019/2020", 
                                ifelse(timeline$siteid=="YELL", 2020,  2019))

timeline$permittingStatus <- ifelse(timeline$siteid %in% c("SJER", "NIWO", "HEAL", "WREF", "BONA"), "site host approved", 
                                    "pending")

timeline$Tower_vst_2019 <- ifelse(timeline$siteid=="NIWO", "scheduled: 9/9 - 9/12", 
                                  ifelse(timeline$siteid=="RMNP", "scheduled: 8/5 - 8/8 (pending site host approval).", 
                                         ifelse(timeline$siteid=="WREF", "scheduled: 8/26 - 9/12.",
                                                ifelse(timeline$siteid%in%c("SOAP", "TEAK"), "characterization data used",
                                                       ifelse(timeline$siteid=="YELL", "inital bout", 
                                                              ifelse(timeline$siteid=="SJER", "", NA)
                                                       )
                                                )
                                         )
                                  )
)

timeline$extend_TFT <- ifelse(timeline$siteid%in%c("BONA", "SJER", "WREF"), "yes",
                              ifelse(timeline$siteid%in%c("NIWO", "HEAL", "RMNP", "DEJU"), "no", NA)
)
                                                       
timeline <- arrange(timeline, proposedYear, desc(permittingStatus), Tower_vst_2019)


panderOptions('table.split.table', Inf)
 pander::pander(timeline, justify="left",  missing='',  caption="Timeline for dendrometer band installation.")

```


# Site Specific Modifications
## RMNP
The site host at this site has requested a maximum of 80 bands be installed. A total of 86 trees qualify for bands according to this analysis. Trees at this site were ordered by biomass and every 14th tree removed to retain the final distribution of tree sizes banded at the site.  

The trees selected for removal include:

* NEON.PLA.D10.RMNP.03039
* NEON.PLA.D10.RMNP.02184
* NEON.PLA.D10.RMNP.02418
* NEON.PLA.D10.RMNP.02311
* NEON.PLA.D10.RMNP.02309
* NEON.PLA.D10.RMNP.02900

