---
title: "HBP D18/D19 2021 Bryophyte Sorting Subsample Experiment"
author: "Courtney L. Meier"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#   Load required libraries
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lattice)
library(lmerTest)
library(openxlsx)

```

##  Goal

To determine the efficacy of subsampling the BRY growth form across various proportions of HBP samples in all D18/D19 sites in order to reduce long sort times associated with bryophytes from current-year clipped biomass. Subsampling is only evaluated for the BRY functional group; all other functional groups are sorted from the full clip strip.



##  Experimental Setup and Analyses

* Subsample the BRY growth from from as many `clipIDs` as possible, in both Distributed and Tower plots.

* For each clipID harvested in the field for which subsampling is performed, create subsamples with the following percentages of the total freshMass:
    - 10%
    - 15%
    - 25%
    - 50%
    - 100% (sum of all subsamples)

* When subsampling is employed, calculate dryMass as follows: `dM = fM * (ssDM/ssFM)`, where:
    - *dM* = estimated dryMass of live BRY in the clipID
    - *fM* = total freshMass of unsorted live and dead BRY in the clipID
    - *ssDM* = subsampleDryMass of live BRY in the subsample
    - *ssFM* = fresh mass of all unsorted live and dead BRY biomass in the subsample (BRY + other clipped material)

* Compare dryMass results calculated via subsampling with dryMass obtained with no subsampling, and use linear mixed-effects models to analyze results.



##  Procedure

1. Perform clip harvest in the field as normal, and bring clipped biomass back to the laboratory in cold storage as normal. For example, sorting of non-BRY functional groups may take place in the field, and sorted non-BRY as well as all clipped BRY (live and dead) are brought back to the lab for further processing.

1. Assuming non-BRY are sorted out, mix unsorted BRY biomass from each clipID to homogenize as thoroughly as possible.
    a. For large amounts of biomass, and when there is more than one bag of biomass for a given clipID, use a large bag, box, tray or equivalent vessel to mix the biomass.

1. For each clipID, weigh and record fresh, unsorted BRY from the entire clip strip to 0.01 g:
    a. **`freshMass`** = total fresh mass in the clipID (all clipped, unsorted BRY material)

1. Based on the **`freshMass`**, calculate the desired subsample fresh masses for testing. For example, assuming **`freshMass`** = 100 g, the target subsample fresh masses are:
    a. 10% subsample --> 10 g
    a. 15% subsample --> 15 g
    a. 25% subsample --> 25 g
    a. 50% subsample --> 50 g

1. Label a coin envelope for each subsample above with the information below.
    a. **`subsampleTest`**: `10%, 15%, 25%` or `50%`.
    a. **`clipID`**
    a. **`collectDate`**

1. Weigh each subsample created above (all clipped BRY material), and record the information below.
    a. **`subsampleTest`**: as above
    a. **`clipID`**
    a. **`collectDate`**
    a. **`subsampleFreshMass`**: To the nearest 0.01 g; for subsamples < 0.5 g total mass, weigh to the nearest 0.0001 g

1. Sort live BRY from other clipped material for each subsample, and place sorted BRY biomass into the corresponding labeled coin envelope. 
    a. Dead BRY, organic material, and other clipped material that is NOT live BRY may be discarded at this point.

1. Dry subsamples until dry; minimum of 48 h @ 65 ˚C, track drying progress as normal.

1. Remove dry samples from the oven one at a time, and immediately weigh and record:
    a. **subsampleDryMass**: To the nearest 0.01 g; for masses < 0.5 g, weigh to the nearest 0.0001 g.

\pagebreak

##  Analyses and Graphs

Using formula from above, `dryMass` values are calculated for BRY from `freshMass` value from whole clip-strip, as well as subsample fresh and dry mass values.

Note: The 15% subsample group has the greatest uncertainty due to the fact it was consistently sorted last and therefore lost the most water.

```{r massCalculations, warning=FALSE, message=FALSE}
# Read in data collected by D18/D19
if(file.exists("/Users/cmeier")){
  inputPath <- "~/Documents/gitRepositories/neon-plant-sampling/hbpSampling/subsampleSortExperiments/"
}

if(file.exists("/Users/Pajaro")){
  inputPath <- "~/Documents/workDocuments/gitRepositories/neon-plant-sampling/hbpSampling/subsampleSortExperiments/"
}

df <- openxlsx::read.xlsx(
  xlsxFile = paste0(inputPath, "hbp_d1819_sortExperiment2021_data.xlsx"),
  sheet = "R_input",
  startRow = 1,
  colNames = TRUE
)


#   Calculate total observed dryMass for each Clip Strip, and estimated
#   dryMass from subsampling
df <- df %>%
  dplyr::mutate(
    dryMass10 = round(freshMass*(ssDM10/ssFM10), digits = 2),
    .before = dryMass
  ) %>%
  dplyr::mutate(
    dryMass15 = round(freshMass*(ssDM15/ssFM15), digits = 2),
    .before = dryMass
  ) %>%
  dplyr::mutate(
    dryMass25 = round(freshMass*(ssDM25/ssFM25), digits = 2),
    .before = dryMass
  ) %>%
  dplyr::mutate(
    dryMass50 = round(freshMass*(ssDM50/ssFM50), digits = 2), 
    .before = dryMass
  ) %>%
  dplyr::arrange(domainID, siteID, clipID)


##  Re-arrange data for lmer models and ggplot2
#   Create long data frame for lmer models
longDF <- df %>%
  dplyr::select(
    domainID,
    siteID,
    clipID,
    collectDate,
    dryMass10,
    dryMass15,
    dryMass25,
    dryMass50,
    dryMass
  ) %>%
  tidyr::pivot_longer(
    cols = tidyr::starts_with("dry"),
    names_to = "treatment",
    values_to = "estimatedDryMass"
  ) %>%
  dplyr::arrange(domainID, siteID, clipID)

#   Create summary data frame for ggplot2
summaryDF <- longDF %>%
  dplyr::group_by(domainID, siteID, treatment) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(estimatedDryMass),
    sd = sd(estimatedDryMass),
    se = sd/sqrt(count)
  )

```

For each site, we graph the true observed BRY dryMass alongside the estimated dryMass calculated via subsampling.

``` {r bryGraphs, echo=FALSE, warning=FALSE, message=FALSE}
### Create graphs of subsampling results by siteID
#   BARR graph
barrGraph <- ggplot2::ggplot(
  data = summaryDF %>%
    dplyr::filter(siteID == "BARR"),
  mapping = aes(
    x = treatment,
    y = mean
  )
) +
  ggplot2::geom_bar(
    position = "dodge",
    stat = "identity"
  ) +
  ggplot2::geom_errorbar(
    aes(
      ymin = mean - se,
      ymax = mean + se
    ),
    width = 0.2,
    position = position_dodge(0.9)
  ) +
  ggplot2::labs(
    x = "Treatment",
    y = "Dry Mass (g)",
    title = "D18 BARR BRY Subsampling"
  )

barrGraph

#   TOOL graph
toolGraph <- ggplot2::ggplot(
  data = summaryDF %>%
    dplyr::filter(siteID == "TOOL"),
  mapping = aes(
    x = treatment,
    y = mean
  )
) +
  ggplot2::geom_bar(
    position = "dodge",
    stat = "identity"
  ) +
  ggplot2::geom_errorbar(
    aes(
      ymin = mean - se,
      ymax = mean + se
    ),
    width = 0.2,
    position = position_dodge(0.9)
  ) +
  ggplot2::labs(
    x = "Treatment",
    y = "Dry Mass (g)",
    title = "D18 TOOL BRY Subsampling"
  )

toolGraph

#   BONA graph
bonaGraph <- ggplot2::ggplot(
  data = summaryDF %>%
    dplyr::filter(siteID == "BONA"),
  mapping = aes(
    x = treatment,
    y = mean
  )
) +
  ggplot2::geom_bar(
    position = "dodge",
    stat = "identity"
  ) +
  ggplot2::geom_errorbar(
    aes(
      ymin = mean - se,
      ymax = mean + se
    ),
    width = 0.2,
    position = position_dodge(0.9)
  ) +
  ggplot2::labs(
    x = "Treatment",
    y = "Dry Mass (g)",
    title = "D19 BONA BRY Subsampling"
  )

bonaGraph

#   DEJU graph
dejuGraph <- ggplot2::ggplot(
  data = summaryDF %>%
    dplyr::filter(siteID == "DEJU"),
  mapping = aes(
    x = treatment,
    y = mean
  )
) +
  ggplot2::geom_bar(
    position = "dodge",
    stat = "identity"
  ) +
  ggplot2::geom_errorbar(
    aes(
      ymin = mean - se,
      ymax = mean + se
    ),
    width = 0.2,
    position = position_dodge(0.9)
  ) +
  ggplot2::labs(
    x = "Treatment",
    y = "Dry Mass (g)",
    title = "D19 DEJU BRY Subsampling"
  )

dejuGraph

#   HEAL graph
healGraph <- ggplot2::ggplot(
  data = summaryDF %>%
    dplyr::filter(siteID == "HEAL"),
  mapping = aes(
    x = treatment,
    y = mean
  )
) +
  ggplot2::geom_bar(
    position = "dodge",
    stat = "identity"
  ) +
  ggplot2::geom_errorbar(
    aes(
      ymin = mean - se,
      ymax = mean + se
    ),
    width = 0.2,
    position = position_dodge(0.9)
  ) +
  ggplot2::labs(
    x = "Treatment",
    y = "Dry Mass (g)",
    title = "D19 HEAL BRY Subsampling"
  )

healGraph

```
\pagebreak


## Results: Mixed-Effects model analysis

- Models are run separately for each site, as the aim is to understand site-specific effects of `treatment` on `estimatedDryMass`. 
- Models use **clipID** as a random effect; benefit of random effect is it can account for variation at level of clip-strip unrelated to treatment - e.g., variation due to moisture on day clip-strip was harvested, effect of moss abundance or species composition at particular location, etc. 
- The version of the `lmer()` function in the `lmerTest` package is used because the assumptions for this model allow standard anova output to be assessed.

```{r barrModel}
### BARR subsampling model
#   Create model with random intercept by clipID
barrMod <- lmerTest::lmer(
  formula = estimatedDryMass ~ treatment + (1|clipID),
  data = longDF %>% dplyr::filter(siteID == "BARR"),
  REML = FALSE
)

#   Assess model outputs
summary(barrMod)
anova(barrMod)
lattice::dotplot(ranef(barrMod, condVar = TRUE))

```
\pagebreak

``` {r toolModel}
### TOOL subsampling model
#   Create TOOL linear mixed-effects model to assess subsampling treatment
toolMod <- lmerTest::lmer(
  formula = estimatedDryMass ~ treatment + (1|clipID),
  data = longDF %>% dplyr::filter(siteID == "TOOL"),
  REML = FALSE
)

#   Assess model outputs
summary(toolMod)
anova(toolMod)
lattice::dotplot(ranef(toolMod, condVar = TRUE))

```
\pagebreak

``` {r bonaMod}
### BONA subsampling model
#   Create BONA linear mixed-effects model to assess subsampling treatment
bonaMod <- lmerTest::lmer(
  formula = estimatedDryMass ~ treatment + (1|clipID),
  data = longDF %>% dplyr::filter(siteID == "BONA"),
  REML = FALSE
)

#   Assess model outputs
summary(bonaMod)
anova(bonaMod)
lattice::dotplot(ranef(bonaMod, condVar = TRUE))

```
\pagebreak

``` {r dejuMod}
### DEJU subsampling model
#   Create DEJU linear mixed-effects model to assess subsampling treatment
dejuMod <- lmerTest::lmer(
  formula = estimatedDryMass ~ treatment + (1|clipID),
  data = longDF %>% dplyr::filter(siteID == "DEJU"),
  REML = FALSE
)

#   Assess model outputs
summary(dejuMod)
anova(dejuMod)
lattice::dotplot(ranef(dejuMod, condVar = TRUE))

```
\pagebreak

``` {r healMod}
### HEAL subsampling model
#   Create HEAL linear mixed-effects model to assess subsampling treatment
healMod <- lmerTest::lmer(
  formula = estimatedDryMass ~ treatment + (1|clipID),
  data = longDF %>% dplyr::filter(siteID == "HEAL"),
  REML = FALSE
)

#   Assess model outputs
summary(healMod)
anova(healMod)
lattice::dotplot(ranef(healMod, condVar = TRUE))

```
\pagebreak




##  Summary

1. BARR results
    - 10% subsampling 14% lower than control, marginally significant (p = 0.078)
    - 15% subsampling 15% lower than control, marginally significant (p = 0.060)
    - 25% subsampling 12% higher than control (p = 0.14)
    - 50% subsampling virtually identical to control (p = 0.85)
1. TOOL results
    - 10% subsampling 6.6% higher than control (p = 0.27)
    - 15% subsampling 4.4% higher than control (p = 0.47)
    - 25% subsampling about 10% lower than control, marginally significant (p = 0.11)
    - 50% subsampling less than 3% higher than control (p = 0.71)
1. BONA results
    - 10% subsampling 3.8% higher than control (p = 0.21)
    - 15% subsampling 3.8% lower than control (p = 0.20)
    - 25% subsampling < 1% lower than control (p = 0.80)
    - 50% subsampling < 1% higher than control (p = 0.78)
1. DEJU results
    - 10% subsampling 11% higher than control (p < 0.05)
    - 15% subsampling 24% lower than control (p < 0.0001)
    - 25% subsampling 2.8% higher than control (p = 0.52)
    - 50% subsampling 3% higher than control (p = 0.49)
1. HEAL results
    - 10% subsampling 2.4% higher than control (p = 0.53)
    - 15% subsampling 4.2% lower than control (p = 0.30)
    - 25% subsampling virtually identical to control (p = 0.997)
    - 50% subsampling virtually identical to control (p = 0.84)
    
Summary of recommendations for BRY subsampling by site:

- BARR: 50%
- TOOL: 50% (conservative due to variable results at lower subsampling levels)
- BONA: 10%
- DEJU: 25%
- HEAL: 10%

Summary of recommendations for BRY subsampling by domain:

- D18 (BARR, TOOL): 50%
- D19 (BONA, DEJU, HEAL): 25%

D18/19 leads indicate it will be easier to train and execute BRY subsampling correctly if a single subsampling target is used per domain.

    