---
title: "Herbaceous Biomass OSD sorting"
author: "Courtney Meier"
date: "10/14/2016"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(xlsx)
library(knitr)

```

## Background
- Sorting OSD from current-year clipped biomass requires an extraordinary amount of time in D09 on a per clip strip basis.
- Kentucky Bluegrass 'mat' is highly intertwined current-year and OSD; like matted hair according to Andrea A. This means that sorting essentially must be done blade by blade in the laboratory following collection in the field.
- Goal here is to determine a data-driven cap on OSD sort time. Because total sort time is quite variable from clip strip to clip strip, would like to know the following:
    - What % of clip/sort events are X% complete in Y minutes?
    - Preliminarily, aim for 95% of plots at 90% complete
- Data source:
    - D09 timed sorting of OSD in 15 or 30 minute intervals for, and weighed total mass of OSD sorted at each interval.
    - Enables analysis of % of cumulative OSD sorted through time on a per plot basis.
    
``` {r dataInput}
if (file.exists(
  '/Users/cmeier/Documents/neonScienceDocs/gitRepositories')){
  wdir <- "~/Documents/neonScienceDocs/gitRepositories/neonPlantSampling/hbpSampling/sampleSortTime"
}

setwd(wdir)

clipSort <- read.xlsx("hbpSampleSorting.xlsx", sheetIndex = 1, header = TRUE)
clipSort <- clipSort %>% select(-NA.) %>% filter(osdMass!="NA")


```

### Data frame manipulation
- Add clipID column
- Add cumulative sum of sorted osdMass and cumulative percent osdMass
``` {r addColumns}
clipID <- paste(clipSort$plotID, clipSort$clipCellNumber, sep = "_")
clipSort$clipID <- clipID

# Calculate cumulative sum and percent of osdMass per clipID, and add to original 'clipSort' dataframe
uniqueClip <- unique(clipSort$clipID)
cumulativeOSD <- NULL
cumulativePercent <- NULL
for (i in 1:length(unique(clipSort$clipID))){
  temp <- filter(clipSort, clipID==uniqueClip[i])
  cumSumTemp <- cumsum(temp$osdMass)
  cumPercentTemp <- round((cumSumTemp/max(cumSumTemp))*100)
  cumulativeOSD <- append(cumulativeOSD, cumSumTemp)
  cumulativePercent <- append(cumulativePercent, cumPercentTemp)
}

clipSort$cumulativeOSD <- cumulativeOSD
clipSort$cumulativePercent <- cumulativePercent

```

### Calculations
- At 30 min time intervals, calculate the % of plots with 90% of sorting complete
``` {r calculations}
theTimes <- seq(from=30, to=330, by=30)

# Calculations for 80% complete OSD sorting
percentPlots <- NULL
plotNum80Complete <- NULL
plotPercent80Complete <- NULL

for (i in 1:length(theTimes)){
  thePlots <- clipSort %>% filter(time <= theTimes[i] & cumulativePercent >= 80)
  numPlots <- length(unique(thePlots$clipID))
  percentPlots <- round((length(unique(thePlots$clipID))/length(uniqueClip))*100, digits = 1)
  plotNum80Complete <- append(plotNum80Complete, numPlots)
  plotPercent80Complete <- append(plotPercent80Complete, percentPlots)
}

# Calculations for 90% complete OSD sorting
percentPlots <- NULL
plotNum90Complete <- NULL
plotPercent90Complete <- NULL

for (i in 1:length(theTimes)){
  thePlots <- clipSort %>% filter(time <= theTimes[i] & cumulativePercent >= 90)
  numPlots <- length(unique(thePlots$clipID))
  percentPlots <- round((length(unique(thePlots$clipID))/length(uniqueClip))*100, digits = 1)
  plotNum90Complete <- append(plotNum90Complete, numPlots)
  plotPercent90Complete <- append(plotPercent90Complete, percentPlots)
}

clipResults <- data_frame(theTimes, plotNum80Complete, plotPercent80Complete, plotNum90Complete, plotPercent90Complete)

```

### Summary of Results
```{r tableSummary}
clipResults %>% kable
```

- blah

