---
title: "HBP D18/D19 2021 Bryophyte Sorting Subsample Experiment"
author: "Courtney L. Meier"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#   Load required libraries
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lmerTest)
library(openxlsx)

```

##  Goal

To determine the efficacy of subsampling the BRY growth form across various proportions of HBP samples in all D18/D19 sites in order to reduce long sort times associated with bryophytes from current-year clipped biomass. Subsampling is only evaluated for the BRY functional group; all other functional groups are sorted from the full clip strip.



##  Experimental Setup and Analyses

* Subsample the BRY growth from from as many `clipIDs` as possible, in both Distributed and Tower plots.

* For each clipID harvested in the field, test procedure by creating subsamples with the following percentages of the total freshMass:
    - 10%
    - 15%
    - 25%
    - 50%
    - 100% (sum of all subsamples)

* When subsampling is employed, calculate dryMass as follows: `dM = fM * (ssDM/ssFM)`, where:
    - *dM* = estimated dryMass of BRY in the clipID
    - *fM* = total freshMass in the clipID
    - *ssDM* = subsampleDryMass of BRY in the subsample
    - *ssFM* = fresh mass of all biomass in the subsample (BRY + other clipped material)

* Compare dryMass results calculated via subsampling with dryMass obtained with no subsampling, and use linear mixed-effects models to analyze results.



##  Procedure

1. Perform clip harvest in the field as normal, and bring clipped biomass back to the laboratory in cold storage as normal.

1. Thoroughly mix biomass from each clipID to homogenize as thoroughly as possible.
    a. For large amounts of biomass, and when there is more than one bag of biomass for a given clipID, use a large bag, box, tray or equivalent vessel to mix the biomass.

1. For each clipID, weigh and record to 0.01 g:
    a. **`freshMass`** = total fresh mass in the clipID (all clipped material)

1. Based on the **`freshMass`**, calculate the desired subsample fresh masses for testing. For example, assuming **`freshMass`** = 100 g, the target subsample fresh masses are:
    a. 10% subsample --> 10 g
    a. 15% subsample --> 15 g
    a. 25% subsample --> 25 g
    a. 50% subsample --> 50 g

1. Label a coin envelope for each subsample above with the information below.
    a. **`subsampleTest`**: `10%, 15%, 25%` or `50%`.
    a. **`clipID`**
    a. **`collectDate`**

1. Weigh each subsample created above (all clipped material), and record the information below.
    a. **`subsampleTest`**: as above
    a. **`clipID`**
    a. **`collectDate`**
    a. **`subsampleFreshMass`**: To the nearest 0.01 g; for subsamples < 0.5 g total mass, weigh to the nearest 0.0001 g

1. Sort BRY from other clipped material for each subsample, and place sorted, current-year BRY biomass into the corresponding labeled coin envelope. 
    a. Sorted OSD may be discarded at this point.

1. Dry subsamples until dry; minimum of 48 h @ 65 ˚C, track drying progress as normal.

1. Remove dry samples from the oven one at a time, and immediately weigh and record:
    a. **subsampleDryMass**: To the nearest 0.01 g; for masses < 0.5 g, weigh to the nearest 0.0001 g.
    


##  Analyses

```{r calculations, echo=FALSE}
# Read in data collected by D09
if(file.exists("/Users/cmeier")){
  inputPath <- "~/Documents/gitRepositories/neonPlantSampling/hbpSampling/subsampleSortExperiments/"
}

df <- read.csv(paste0(inputPath, "hbp_d17SortExperiment2017_input.csv"), header = TRUE, stringsAsFactors = FALSE)

#   Calculate total observed dryMass for each Clip Strip, and estimated dryMass from subsampling
#   Filter to focus analyses on 2018 bout collected between 2/19-2/28; concern that earlier bout did not 
#   adequately represent clipping/sorting conditions that exist during majority of field season.
df %>%
  mutate(dryMass = ssDM10 + ssDM25 + ssDM50 + ssDMResid) %>%
  mutate(dryMass10 = freshMass*(ssDM10/ssFM10)) %>%
  mutate(dryMass25 = freshMass*(ssDM25/ssFM25)) %>%
  mutate(dryMass50 = freshMass*(ssDM50/ssFM50)) %>%
  filter(grepl("2018", eventID)) %>%
  arrange(eventID, clipID, exclosure) -> df

# Arrange data suitable for ggplot2, then plot
df %>%
  select(eventID, clipID, collectDate, exclosure, dryMass, dryMass10, dryMass25, dryMass50) %>%
  gather(key = treatment, value = estimatedDryMass, dryMass10, dryMass25, dryMass50, dryMass) %>%
  rename() %>%
  arrange(clipID, treatment) -> longDF

summaryDF <- ddply(longDF, c("treatment", "exclosure"), summarise,
                   N = length(estimatedDryMass),
                   mean = mean(estimatedDryMass),
                   sd = sd(estimatedDryMass),
                   se = sd/sqrt(N))

ggplot(summaryDF, aes(x = treatment, y = mean, fill = exclosure)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "Treatment", y = "Dry mass (g)", title = "D17 HBP Clip Strip subsampling")


```

### Results: Mixed-Effects model analysis

1.  Model accounting for subsampling effect on **dryMass** and using **clipID** as a random effect (no need to use **exclosure** as a fixed effect since random effect accounts for variation across exclosure treatment).
```{r meTreatment}

m1ML <- lmer(estimatedDryMass ~ treatment + (1|clipID), longDF, REML = FALSE)
summary(m1ML)
anova(m1ML)

```
\pagebreak


2.  Null model, and using **clipID** as a random effect.
```{r meExclosure}
m3ML <- lmer(estimatedDryMass ~ 1 + (1|clipID), longDF, REML = FALSE)
summary(m3ML)
anova(m1ML, m3ML)

```
\pagebreak


3.  Linear model, using treatment*exclosure, and no random effect.
```{r lmExclosureInterxnTreatment}
m1LM <- lm(estimatedDryMass ~ exclosure*treatment, data = longDF)
summary(m1LM)
anova(m1LM)

```
\pagebreak


##  Outcomes

1. Random effect associated with **clipID** explains considerable variability in the data. See summary output for model = m1ML.
1. Summary output for Mixed Effects model indicates that D17 Clip Strips should be subsampled at the 25% level for all bouts with **herbGroup** = ALL.
    - Estimates of dryMass using a 10% subsample are significantly higher than the entire sorted sample, using data from one bout collected late February and just before peak green.
    - Other subsamples are not significantly different than the entire sorted subsample --> go with 25%
1. Effect of `exclosure*treatment`, as evaluated in standard linear model:
    - Effect of `exclosure` not significant (m1LM), and effect of `treatment` also not significant. No detection of `exclosure:treatment` interaction effect.

    