---
title: "HBP D11 2023 Sorting Subsample Experiment"
author: "Courtney Meier"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Goal
To determine the efficacy of subsampling various proportions of HBP samples in D11 to reduce long sort times associated with removing OSD from current-year clipped biomass. **Subsampling is only evaluated in the context of clip harvests that do not require sorting to functional group** (i.e., non-peak biomass clips), because the subsample to total mass ratios will not apply to individual herbGroups.

##  Experimental Setup and Analyses

* Select n=10 plots (all plots for OAES), resulting in a maximum of n=20 `clipID` to test, due to both `exclosure = Y and N` for each plot. Random spatially-balanced plot locations, and locations of clipIDs within plots, will provide an unbiased estimate of biomass throughout the Tower airshed.

* For each `clipID` harvested in the field, test subsampling efficiency at various levels of sorting by creating subsamples (current-year + OSD) with the following percentages of the total `freshMass`:
    - 10%
    - 15%
    - 25%
    - 50%
    
    The sum of all the subsamples = 100%; that is, the fresh mass of the entire clip strip.

* When subsampling is employed, calculate dryMass as follows: `dM = fM * (ssDM/ssFM)`, where:
    - *dM* = dryMass of current-year biomass in the clipID (no OSD)
    - *fM* = total freshMass in the clipID (current-year + OSD)
    - *ssDM* = subsampleDryMass of current-year biomass in the subsample (no OSD)
    - *ssFM* = fresh mass of all biomass in the subsample (current-year + OSD)

* Compare `dryMass` results calculated via subsampling with `dryMass` obtained with no subsampling, and use mixed effects models to analyze results.

##  Procedure

1. Perform clip harvest in the field as normal, and bring clipped biomass back to the laboratory in cold storage as normal.

1. Identify up to n=20 clipIDs (n=10 for `exclosure = Y` and n=10 for `exclosure = N`), originating from each of the 10 plotIDs.

1. Thoroughly mix biomass from each `clipID` to homogenize as thoroughly as possible.
    a. For large amounts of biomass, and when there is more than one bag of biomass for a given `clipID`, use a large bag, box, tray or equivalent vessel to mix the biomass.

1. For each clipID, weigh and record to 0.01 g:
    a. **`freshMass`** = total fresh mass in the clipID (current-year + OSD).
    a. It is important that `freshMass` for the entire `clipID` and subsample fresh masses (below) are collected for a given `clipID` as close to each other in time as possible. That is, avoid weighing `freshMass` for a given `clipID` hours apart from the subsample fresh masses as water loss will affect the experimental results.

1. Based on the **`freshMass`**, calculate the desired subsample fresh masses for testing. For example, assuming **`freshMass`** = 100 g, the target subsample fresh masses are:
    a. 10% subsample --> 10 g
    a. 15% subsample --> 15 g
    a. 25% subsample --> 25 g
    a. 50% subsample --> 50 g

1. Label a coin envelope for each subsample above with the information below.
    a. **`subsampleTest`**: `10%, 15%, 25%` or `50%`
    a. **`clipID`**
    a. **`collectDate`**
    a. **`exclosure`**: `Y/N`

1. Weigh each fresh subsample created above (current-year + OSD), and record the information below.
    a. **`subsampleTest`**: as above
    a. **`clipID`**
    a. **`collectDate`**
    a. **`exclosure`**: `Y/N`
    a. **`subsampleFreshMass`**: To the nearest 0.01 g; for subsamples < 0.5 g total mass, weigh to the nearest 0.0001 g

1. Sort current-year biomass from OSD for each subsample, and place sorted, current-year biomass into the corresponding labeled coin envelope.
    a. Sorted OSD may be discarded at this point.

1. Dry subsamples until dry; minimum of 48 h @ 65 ˚C, track drying progress as normal.

1. Remove dry samples from the oven one at a time, and immediately weigh and record:
    a. **subsampleDryMass**: To the nearest 0.01 g; for masses < 0.5 g, weigh to the nearest 0.0001 g.
    
    
##  Analyses

```{r loadPackages, echo=FALSE, include=FALSE}

library(dplyr)
library(ggplot2)
library(glue)
library(lmerTest)
library(openxlsx)
library(tidyr)

```
    
```{r calculations, echo=FALSE}

### Read in data collected by D11
if(file.exists("/Users/cmeier")){
  inputPath <- "~/Documents/gitRepositories/neon-plant-sampling/hbpSampling/subsampleSortExperiments"
}

df <- openxlsx::read.xlsx(xlsxFile = glue::glue(inputPath, "hbp_d11SortExperiment2023_data.xlsx", .sep = "/"),
                          sheet = "R_input",
                          colNames = TRUE) %>%
  dplyr::mutate(collectDate = lubridate::ymd(collectDate))



### Calculate total dryMass for each sampling cell, and estimated dryMass from subsampling
df <- df %>%
  dplyr::mutate(dryMass10 = round(freshMass * (ssDM10/ssFM10), digits = 2),
                dryMass15 = round(freshMass * (ssDM15/ssFM15), digits = 2),
                dryMass25 = round(freshMass * (ssDM25/ssFM25), digits = 2),
                dryMass50 = round(freshMass * (ssDM50/ssFM50), digits = 2),
                dryMass = ssDM10 + ssDM15 + ssDM25 + ssDM50) %>%
  dplyr::arrange(eventID, clipID, exclosure)



### Create long-format and summary data frames for LMER models and ggplot2

#   Generate long-format data frame
longDF <- tidyr::pivot_longer(data = df %>%
                                dplyr::select(eventID, clipID, collectDate, exclosure,
                                              dryMass10, dryMass15, dryMass25, dryMass50, dryMass),
                              cols = dryMass10:dryMass,
                              names_to = "treatment",
                              values_to = "estDryMass")



#   Calculate total observed dryMass for each Clip Strip, and estimated dryMass from subsampling
#   Filter to focus analyses on 2018 bout collected between 2/19-2/28; concern that earlier bout did not 
#   adequately represent clipping/sorting conditions that exist during majority of field season.
df %>%
  mutate(dryMass = ssDM10 + ssDM25 + ssDM50 + ssDMResid) %>%
  mutate(dryMass10 = freshMass*(ssDM10/ssFM10)) %>%
  mutate(dryMass25 = freshMass*(ssDM25/ssFM25)) %>%
  mutate(dryMass50 = freshMass*(ssDM50/ssFM50)) %>%
  filter(grepl("2018", eventID)) %>%
  arrange(eventID, clipID, exclosure) -> df

# Arrange data suitable for ggplot2, then plot
df %>%
  select(eventID, clipID, collectDate, exclosure, dryMass, dryMass10, dryMass25, dryMass50) %>%
  gather(key = treatment, value = estimatedDryMass, dryMass10, dryMass25, dryMass50, dryMass) %>%
  rename() %>%
  arrange(clipID, treatment) -> longDF

summaryDF <- ddply(longDF, c("treatment", "exclosure"), summarise,
                   N = length(estimatedDryMass),
                   mean = mean(estimatedDryMass),
                   sd = sd(estimatedDryMass),
                   se = sd/sqrt(N))

ggplot(summaryDF, aes(x = treatment, y = mean, fill = exclosure)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "Treatment", y = "Dry mass (g)", title = "D17 HBP Clip Strip subsampling")


```
