---
title: "EAB quarantine"
author: "FSU"
date: \today
output: pdf_document
---

# Summary of analyses

An Observatory wide approach to addressing sites under quarantine for Emerald Ash Borer is still under development; until guidance is available, implementation of the Litterfall and Fine Woody Debris protocol (NEON.DOC.001710) will only occur in plots where \textit{Fraxinus sp.} is not present. Data for summary tables of \textit{Fraxinus sp.} presence/absence comes from the vegetation characterization effort and implementation of the vegetation structure protocol. Though a plot may indicate that it is safe to collect material, this is intended to be general guidance; if Ash species are present in the plot, please discontinue litter sampling. 


```{r loadLibraries, echo=F, message=F}


library(plyr)
library(dplyr)
library(knitr)
library(stringr)
library(readr)

```

```{r filePaths, echo=F, message=F}

if(file.exists(
  'C:/Users/kjones')){
  myPathToData<- 'C:/Users/kjones/SharePoint/Veg. Characterization - Documents/Data/Output CSVs/ready for dodobase'
  myPathToRepo <- 'C:/Users/kjones/Documents/GitHub/neonPlantSampling/EAB_quarantine'
  myPathToDev <- 'C:/Users/kjones/Documents/GitHub/devTOS'
  myPathtoIPTrepo<-'C:/Users/kjones/Documents/GitHub/organismalIPT/'
}

# taxFile<-'N:/common/TOS/ATBDs/Lookups/pla_names_status_list3.csv'
# taxTable<-read.table(taxFile, na.strings='', sep=',', fileEncoding='UTF-8')
# titles<-read.csv(paste(myPathtoIPTrepo, 
#                        'scienceDev/lookupTablesShared/taxonomyTemplate.csv', sep=''))
# names(taxTable)<-titles$fieldName[1:95]
#

species <- read.csv(paste(myPathToRepo, 'fraxSpecies.csv', sep='/'), stringsAsFactors = F)
for (i in 1:nrow(species)){
  if (species$taxonID[i]==""){
  species$taxonID[i] <- species$acceptedTaxonID[i]
  }
}
species <- select(species, taxonID, scientificName)

plotData <- read.csv(paste(myPathToRepo, 'mOrder_plots.csv', sep='/'), stringsAsFactors = F)
plotType <- select(plotData, siteID, plotID, plotType)

nsTrees <- read.csv('N:/Science/FSU/SiteCharacterization/2013 Site Characterization/2013_siteChar_veg data/NatureServe/cleaned by CSP 20150504/site_characterization_trees.csv', stringsAsFactors=F)


```


```{r mergeDataFunctions, message=F, echo=F}
##### copied from ~GitHub\codingSupportGroup\FSUTools\file_merging\pull_all_filepaths_in_folderDirectory.R ##############

multipleCombine <- function(input, ply = llply){
  ply(input, function(x){
    t <- read.csv(x, header=TRUE, sep=",",stringsAsFactors = FALSE) # read the csv
    t1 <- rbind(t) # rbind it to a temporary variable
    return(t1) # return the full variable
  }
  )
}

# a re-usable function - outputType can be 'vector' or 'list'
fileParser <- function(directoryList, outputType = "vector"){
  # need to dive into each folder, decide if it's a .csv or .xlsx, then read the rows
  fullFilenameList <- list() # initialize a list object to populate
  counter = 0
  for (folder in directoryList){
    # browser() # this is the interactive debugger, its scope is global & local
    # list all files in a particular folder
    folder_content <- list.files(folder, full.names = TRUE)
    for (fileN in folder_content){
      # iterate the file count for tracking and for indexing the list
      counter = counter + 1
      fullFilenameList[[counter]] <- fileN
      print(fileN)
    }
  }
  # what kind of output should be returned?
  ifelse(outputType == "vector", 
         # return a vector
         return(unlist(fullFilenameList)),
         # return a list
         return(fullFilenameList))
}
##

```

```{r charDataMerge, echo=F, message=F}


directory <- myPathToData

#list all files
folderList <- list.files(directory, full.names = TRUE)

# grep VST files 
vstList2 <- grep(pattern = "VST|vst", x = folderList, value = TRUE)

# grep VST mapping and tagging only
mappingList3 <- grep(pattern = "mapping", x = vstList2, value = TRUE)

# grep .csv 
mappingList4 <- grep(pattern = ".csv", x = mappingList3, value = TRUE)

# save the stack
char <- multipleCombine(mappingList4, ply = ldply)
char <- left_join(char, species)
char <- select(char, siteID, plotID, taxonID, scientificName)

```

```{r vstData, message=F, echo=F}

pre2016 <- read.csv(paste(myPathToDev, 'fulcrum/fulcrum_table_data/vst_lookups/pre2016_dataMerge_mapping_apparentInds/vst_mapping_tagging_L0.csv', sep='/'), stringsAsFactors = F) 

pre2016 <- select(pre2016, -scientificName)
pre2016 <- left_join(pre2016, species)
pre2016 <- select(pre2016, siteID, plotID, taxonID, scientificName)

unde <- filter(nsTrees, Site_abbrev=='UNDE')
#unde data on dropbox contains distributed plots only, must use NatureServe data and reassign plot_IDs
IDmatch <- select(plotData, mortonOrder, siteID, plotID)
IDmatch$mortonOrder <-   as.character(IDmatch$mortonOrder)
unde <- select(unde, siteID=Site_abbrev, mortonOrder=Plot, taxonID=Taxon_Symbol, scientificName=Scientific.Name)
unde <- left_join(unde, IDmatch)


ornl <- read.csv('Z:/2015data/D07/VegStructure/VST_ORNL_2015/vst_ORNL_mapping_in_2014_2015.csv', stringsAsFactors=F)

ornl$siteID <- toupper(substr(ornl$plotID, 1, 4))
ornl$taxonID <- NA
ornl <- select(ornl, siteID, plotID, taxonID, scientificName)

new <- bind_rows(pre2016, char, unde, ornl)
tower <- filter(plotType, plotType=='tower')
new <- left_join(new, tower)
new <- filter(new, !is.na(plotType))

```



```{r function, echo=FALSE, message=FALSE}

fraxCheck <- function(ID=ALL, df=NULL){
    df <- tbl_df(df)
    # sub <- df%>%
    #   select(siteID, plotID, taxonID, scientificName)
    sub <- df[!duplicated(df), ]
 #error if siteID is not contained in specified df
    if(!(ID %in% unique(df$siteID))){
      message(paste("df does not contain tower plot data for", ID))
    }else{
   sub2 <- filter(sub, siteID==ID)
  #empty dataFrame to hold results
  frax <- data.frame(plotID=unique(sub2$plotID),
                    fraxPresent= NA) 
  #remove rows with invalid plotIDs
  frax <- filter(frax, grepl('_', plotID))
  for(i in 1:nrow(frax)){
    frax$fraxPresent[i] <- ifelse(any(grepl("Fraxinus", sub2$scientificName[frax$plotID[i]])), "yes", "no")
  }
  frax <- frax[!duplicated(frax), ]
  frax <- frax %>% 
    arrange(plotID)

   if(all(frax$fraxPresent=='no')){
     message("According to available data, there are no tower plots at ", ID, " that contain Fraxinus sp.", sep=' ')
     message("Litterfall sampling may occur in all plots as directed in the protocol.")
   }else{
#     frax2 <- filter(frax, fraxPresent=='yes')
    #print output
 kable(frax, format="markdown", digits=1)
   }
   }
}


```

## D01 - HARV



```{r HARV, echo=F, comment=NA}

fraxCheck(ID="HARV", df=new)

```


##D01 - BART
```{r BART, echo=F, comment=NA}

fraxCheck(ID="BART", df=new)

```



##D02 - SCBI
```{r SCBI, echo=F, comment=NA}

fraxCheck(ID="SCBI", df=new)

```



##D02 - SERC
```{r SERC, echo=F, comment=NA}

fraxCheck(ID="SERC", df=new)

```


##D05 - UNDE

```{r UNDE, echo=F, comment=NA}

fraxCheck(ID="UNDE", df=new)

```


##D05 - STEI

```{r STEI, echo=F, comment=NA}

fraxCheck(ID="STEI", df=new)

```


##D05 - TREE

```{r TREE, echo=F, comment=NA}

fraxCheck(ID="TREE", df=new)

```


##D06 - UKFS

```{r UKFS, echo=F, comment=NA}

fraxCheck(ID="UKFS", df=new)

```

##D06 - KONZ

```{r KONZ, echo=F, comment=NA}

fraxCheck(ID="KONZ", df=new)

```


##D07 - ORNL

```{r ORNL, echo=F, comment=NA}

data <- read.csv('Z:/2015data/D07/VegStructure/VST_ORNL_2015/vst_ORNL_mapping_in_2014_2015.csv', stringsAsFactors=F)

data$siteID <- toupper(substr(data$plotID, 1, 4))
data$taxonID <- NA

fraxCheck(ID="ORNL", df=new)

```

##D07 - GRSM

```{r GRSM, echo=F, comment=NA}

fraxCheck(ID="GRSM", df=new)

```

##D07 - MLBS

```{r MLBS, echo=F, comment=NA}

fraxCheck(ID="MLBS", df=new)

```