---
title: "D20 Diameter Method Test"
author: "Courtney Meier"
date: "2020-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required libraries
library(dplyr)
library(knitr)
library(PairedData)

# Read in data
df <- read.csv(file = "PUUM_diamMethodTest.csv", header = TRUE, stringsAsFactors = FALSE)

```

### Stem Diameter method evaluation
A subset of large Ohi'a trees at the NEON D20 PUUM site are supported by large prop roots, creating a bole that becomes regular far higher than 130 cm above the ground. To determine whether a pole-based tape-wrapping technique or calipers might be suitable for reaching measurement locations that cannot be reached from the ground with a standard DBH tape approach, staff collected stem diameter data from a sample of trees (n=20) that could be reached from the ground using three techniques:

- Standard DBH tape (control)
- DBH tape + pole
- Calipers

Here, we use two paired t-tests to determine whether the pole-based method and the caliper method are significantly different from the DBH tape control method.

Summary of input data:
```{r diamData, echo=FALSE, message=FALSE, warning=FALSE}
datSum <- dplyr::group_by(df, method) %>%
  summarise(
    count = n(),
    mean = mean(diameter, na.rm = TRUE),
    stdev = sd(diameter, na.rm = TRUE)
  )

kable(datSum)

```


#### Compare Stem Diameter: Tape control vs. Pole method
```{r poleCompare, echo=FALSE}
# Filter desired data
dat <- df %>% filter(method %in% c("dbhTape", "pole"))
dbhTape <- dat[dat$method=="dbhTape", "diameter"]
pole <- dat[dat$method=="pole", "diameter"]

# Conduct t-test
t.test(dbhTape, pole, paired = TRUE, conf.level = 0.95)

# Create paired plot
pd <- paired(dbhTape, pole)
plot(pd, type = "profile")

```

- Paired t-test: No significant difference between control DBH tape and pole-based method.
- Average difference in stem diameter was 0.04 cm larger with pole-based method.
- Difference of > 1 cm apparent for at least 1 individual.


#### Compare Stem Diameter: Tape control vs. Caliper method
```{r caliperCompare, echo=FALSE}
# Filter desired data
dat <- df %>% filter(method %in% c("dbhTape", "caliper"))
dbhTape <- dat[dat$method=="dbhTape", "diameter"]
caliper <- dat[dat$method=="caliper", "diameter"]

# Conduct t-test
t.test(dbhTape, caliper, paired = TRUE, conf.level = 0.95)

# Create paired plot
pd <- paired(dbhTape, caliper)
plot(pd, type = "profile")

```

- Paired t-test: No significant difference between control DBH tape and caliper method.
- Average difference in stem diameter was 0.05 cm larger with caliper method.
- Although stem diameter not significantly higher with caliper method, variation in paired differences is greater compared to the pole method.


#### Conclusions
- Pole-wrapping method is suitable for large Ohi'a trees when it is not possible to reach the desired measurement location from the ground.
- The caliper method may be suitable for difficult individuals that cannot be safely measured from the ground due to terrain, but the caliper method is not preferred.
- It may be possible to improve the caliper method by making multiple diameter measurements - e.g., two measurements orthogonal to each other. Here, only one diameter measurement was collected with the calipers.
