---
title: "CDW maximum probable diameter by site"
author:
- Courtney Meier, 
- Sam Simkin
date:
- 09 September 2019, 
- rev. 29 Apr 2024
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load required libraries
library(tidyverse)
library(knitr)


# Load required functions

source("cdwDlimAndRDFunctions.R")

if (file.exists("/Users/cmeier")){
  source("~/Documents/gitRepositories/neonPlantSampling/cdw_transectLengthAnalysis/cdwDlimAndRDFunctions.R")
}

vst <- neonUtilities::loadByProduct(dpID = "DP1.10098.001",
                                    site = "YELL",
                                    startdate = "2012-01",
                                    enddate = "2023-12",
                                    package = "basic",
                                    release = "LATEST",
                                    tabl = "all",
                                    include.provisional = TRUE,
                                    check.size = FALSE,
                                    token = Sys.getenv("NEON_PAT"))



##  Extract tables from list to global environment
list2env(vst, envir = .GlobalEnv)

plotType <- vst_perplotperyear %>% select(plotID, plotType) %>% distinct(plotID, plotType)
```

# Background
- Field Operations technicians requested a standardized procedure for determining the length of transects for CDW tally sampling. To support this request, I have analyzed Vegetation Structure / Vegetation Characterization DBH data on a per domain basis. 
- For most sites, datasets are filtered to remove stems with DBH <= 10cm; filtered stem DBH data are then analyzed visually with histogram plots, and the 95th percentile of DBH was also calculated. The initial assumption is that the 95th percentile of DBH will correspond to the maximum likely DBH encountered.
- The choice of F-value for each site is initially based on the maximum expected size of the particles, as well as qualitative estimates of particle density at the site. The F-value may be revised iteratively with feedback from Field Operations technicians.
- Results are documented in NEON-2474 (https://neoninc.atlassian.net/browse/NEON-2474).


##  D12 YELL Plots: Pull VST data from L0 database
```{r d12YellDistLoadData, eval=FALSE, include=FALSE}

# Define working filepath
lidsPath <- "C:/GitHub/neonPlantSampling/cdw_transectLengthAnalysis"

# Get VST Apparent Individual data
vst_mappingandtagging <- vst_mappingandtagging %>% select(individualID, taxonID)

yellVst <- merge(vst_apparentindividual, vst_mappingandtagging, by = "individualID", all.x = TRUE)
yellVst <- merge(yellVst, plotType, by = "plotID", all.x = TRUE)
yellVst <- yellVst %>% filter(stemDiameter!="NA") %>%
  select(siteID, eventID, date, plotID, plotType, individualID, taxonID, growthForm, stemDiameter, basalStemDiameter, plantStatus)

# Check per plot sampling effort
yellVst %>% count(plotID) #%>% kable
length(unique(yellVst$plotID))

```

``` {r d12YellDistDupCheck, eval=FALSE, include=FALSE}
# Check for and remove duplicates in input dataframe using dplyr::distinct()
yellDupNum <- nrow(yellVst)-nrow(dplyr::distinct(yellVst)) # -> No duplicates

# Determine number of likely bouts in which data were collected
dates <- sort(unique(yellVst$date))

# Write out data used for analysis
write.csv(dplyr::distinct(yellVst), file = paste("vst_D12_YELL_cleanData.csv", sep = "/"), fileEncoding = "UTF-8", row.names = FALSE)

```

**_Summary of YELL Plot L0 Data_**

- 1640 records in L0 for which a **stemDiameter** was recorded (578 after filtering to the most recent eventID for each plotID)
- 407 records for individuals with DBH >= 10 cm



```{r yellDistDBHAnalysis, include=FALSE}
domainID <- "D12 D"
siteID <- "YELL"
fVal <- 8
# if (file.exists("/Users/cmeier")){
#   inputPath <- "~/Documents/gitRepositories/neonPlantSampling/cdw_transectLengthAnalysis"
# }
# 
 yellVst <- read.csv(paste("vst_D12_YELL_cleanData.csv", sep = "/"), header=T, stringsAsFactors = F)

# YELL: Check input data to determine whether trees with DBH >=10 cm exist in dataset
hist(yellVst$stemDiameter, breaks = 30) # -> Most trees have DBH >= 10 cm, so filter out stems < 10 cm diameter as for other sites

plot_event <- yellVst %>% distinct(plotID, eventID) %>% group_by(plotID) %>% top_n(1, eventID) # identify unique plot by event combos and retain only the latest eventID for each plotID
   
tempVst <- merge(yellVst, plot_event, by = c("plotID","eventID"), all.y = TRUE) # filter to just the most recent eventID for each plotID

tempVst  %>% filter(stemDiameter >=10) -> tempVst
#yellVst %>% filter(stemDiameter >=10 & (eventID == "vst_YELL_2018" | eventID == "vst_YELL_2019")) -> tempVst
#yellVst %>% filter(stemDiameter >=10 & (eventID == "vst_YELL_2021" | eventID == "vst_YELL_2022" | eventID == "vst_YELL_2023")) -> tempVst
#yellVst %>% filter(stemDiameter >=10 & (eventID == "vst_YELL_2020" | eventID == "vst_YELL_2021")) -> tempVst


# Calculate maxLikelystemDiameter, and maxstemDiameter, rounding up to nearest cm
count <- nrow(tempVst)
maxLikelySD <- ceiling(quantile(tempVst$stemDiameter, probs=c(0.95), na.rm = TRUE))[[1]]
 tempVst %>% group_by(eventID) %>% summarise( maxLikelyStd = ceiling(quantile(stemDiameter, probs=c(0.95) )), count = n())

maxSD <- max(tempVst$stemDiameter, na.rm = TRUE)
limDist <- signif(dlim(maxLikelySD, 3, fVal), digits=2)

# Add calculated parameters to existing dlim.df
dlim.df <- data.frame(domainID, siteID, fVal, count, maxLikelySD, maxSD, limDist)
colnames(dlim.df) <- c("domain", "site", "fValue", "count", "maxLikelyStemDiam", "maxStemDiam", "Dlim")

```


**_YELL: Results of stem diameter analysis _**

``` {r d12YellDistHistPlot, echo=FALSE} 
titleText <- paste(domainID, siteID, "YELL stemDiameter distribution", sep=" ")
ggplot(tempVst, aes(x=stemDiameter)) +
  geom_histogram(binwidth=1) +
  geom_vline(xintercept = maxLikelySD, col="blue") +
  labs(title=titleText, x="stemDiameter (cm)", y="Count")

```

- maxLikelystemDiameter = 95th percentile for `stemDiameter` --> `r maxLikelySD` cm
- Max `stemDiameter` = `r maxSD` cm
- Dlim for maxLikelyStemDiameter = `r limDist` m @ F=`r fVal`






\pagebreak

# Summary
##  CDW F-values and transect length (Dlim) by site
Desired transect lengths (`Dlim`, meters) associated with the `maxLikelyStemDiameter` (cm) at each site. The maximum likely stemDiameter is calculated as the 95th percentile for stemDiameter as recorded in Vegetation Structure data from both Distributed and Tower Plots in the most recent eventID for each plot, and `count` indicates the total number of individuals with DBH >= 10 cm used for the analysis; exception(s) are HEAL, where individuals with DBH >= 2 cm were used for the analysis. Transect lengths are based on F-values specific to each site, and will change if a new F-value is chosen for a site.
``` {r tableOutput, echo=FALSE}
dlim.df
```
