---
title: "Optimizing Tower Plot number in AK with NIWO HBP data"
author: "Courtney Meier"
date: "19 April 2016"
output: pdf_document
---

### Load parsed L0 data from Viewer
Data were parsed using the DPS Shiny App (from Claire L.)
https://cklunch.shinyapps.io/parser_app/

Need to fix some residual errors in parsed data, as following variables did not get populated by parser: 

* domainID
* siteID
* plotID
* subplotID <- delete this field, not needed for analysis

In addition:

* Date field requires simplifying to take data left of 'T' character
* Filter out records with `qaDryMass=Y`
```{r loadLibraries, echo=FALSE, include=FALSE}
# Load required libraries
library(dplyr)
library(stringr)

```

```{r loadData}
# Read in output file from DPS parser
parsed.df <- read.csv("NIWO_hbpL0ParsedData_2015_v1.csv", header=T, stringsAsFactors = F)

# Add domainID, siteID, and plotID; remove subplotID
parsed.df$domainID <- "D13"
parsed.df$siteID <- "NIWO"
parsed.df$plotID <- str_sub(parsed.df$clipID, 1, 8)
parsed.df <- select(parsed.df, -subplotID, -remarks)

# Simplify `date` field
#  Creates a list with one list item per 'date', each list item has two elements one for data left of 'T', another for data right of 'T'
date <- str_split(parsed.df$date, "T")
# Here, the x[1] specifies we keep the first column of the dataframe created by ldply; to keep all columns, use `x[]`
date <- plyr::ldply(date, function(x) x[1])
# The df created by ldply has column names V1, V2, etc., and don't want that name, so need to rename
parsed.df$date <- date$V1

# Filter out values with `qaDryMass=Y`
parsed.df <- filter(parsed.df, qaDryMass=="N")

# Write out filtered, cleaned parsed.df to a .csv file
write.csv(parsed.df, file="NIWO_hbpL0ParsedData_2015_v2.csv", row.names=F)


```

### Check for NAs
Identify records with plotID, clipID, or date that have `NA` values for any of these three fields
```{r naCheck}
# The dplyr::tbl_df function makes display of the data frame in the console much more reasonable.
niwoHbp <- tbl_df(read.csv("NIWO_hbpL0ParsedData_2015_v2.csv", header=T, stringsAsFactors = F))

# Identify incomplete records in the dataframe
which(is.na(niwoHbp$plotID))
# -> rows 78, 116, 124

which(is.na(niwoHbp$clipID))
# -> rows 78, 116, 124

which(is.na(niwoHbp$date))
# -> All are complete

# Next, need to fix the 'NA' values identified
niwoHbp$plotID[78] <- "NIWO_054"
niwoHbp$clipID[78] <- "NIWO_054_0186"

niwoHbp$plotID[116] <- "NIWO_046"
niwoHbp$clipID[116] <- "NIWO_046_0091"

niwoHbp$plotID[124] <- "NIWO_046"
niwoHbp$clipID[124] <- "NIWO_046_0091"

```


### Calculate `totalDryMass` for Tower Plots only
Simulated sample sizes will require total `dryMass` values for a given `clipID`. For NIWO, there are no exclosures, and there should only be one clipID per plot. However, need to verify this assumption.
```{r totalDryMass}
# Group the data, calculate the sum of dryMass values and then store as a new output
niwoHbp %>%
  group_by(plotID, date, clipID) %>%
  summarise(totalDryMass = sum(dryMass, na.rm=TRUE)) %>%
  arrange(plotID) -> niwoTot

# Check for duplicates in plotID and clipID
niwoTot %>%
  group_by(plotID, clipID) %>%
  summarize(n=n()) -> dupNiwo

# Join data in niwoTot with applicableModules.csv based on plotID and plotType - want to determine which plotIDs in niwoTot are Tower Plots
setwd("~/Documents/neonScienceDocs/gitRepositories/devTOS/spatialData/supportingDocs")
am <- read.csv("applicableModules.csv", header=T, stringsAsFactors = F)
am %>% filter(siteID=="NIWO", subtype=="basePlot") %>%
  select(plotID, plotType, subtype) -> amNiwo

# Use a left join to keep all rows of niwoTot; plotIDs in niwoTot with no match in amNiwo will be returned with NAs for the columns in amNiwo
niwoTot <- left_join(niwoTot, amNiwo, by="plotID")
niwoTot %>% filter(plotType=="tower") %>%
  select(-subtype) -> niwoTot

# Verify number of records is not > 30 for Tower Plots
nrow(niwoTot)   # -> 29; good
  
# Write out as .csv
write.csv(niwoTot, file="hbp_niwoTotDryMass_2015.csv", row.names = F)

```

Based on results above, `totalDryMass` data from Tower Plots are ready to use for the sampling simulation as there are no duplicates based on `plotID` and `clipID`.


### Sampling simulation
* Goal is to sample with replacement at a variety of reduced sample sizes, calculate a mean for each sampling event, and create a distribution of means at each sample size. 
* The 95% CI for the original n=29 sample size will also be bootstrapped, and for each sample size, the % of means falling within the original 95% CI of the data will be calculated.
* A priori, desire to know the smallest sample size at which ≥ 90% of sampled means fall within the 95% CI intervals of the original data.
``` {r simulation}
# Calculate the 95% CI for the original data (n=29 Tower Plots)


```

``` {r exampleBootstrap}
# Consider the log-normal distibution
x <- rlnorm(1000)
plot(density(x), xlim=c(0, 10))
abline(v=mean(x), col="red")
abline(v=mean(x, tr=.20), col="darkgreen")
abline(v=median(x), col="blue")

library(simpleboot)
# 20% trimmed mean bootstrap
b1 <- one.boot(x, mean, R=2000, tr=.2)
boot.ci(b1, type=c("perc", "bca"))
# The bootstrap trimmed mean:
b1$t0

```

