---
title: "BBC 2023 Alaska Subsample Experiment"
author: "Courtney Meier"
date: "2024-06-26"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##  Goal
To determine the efficacy of subsampling BBC core samples at the D18 TOOL and D19 DEJU sites to reduce time required to sort roots into size categories. 


##  Experimental Setup and Analyses

* Select n=10 cores for each plot site. Cores are sampled from random spatially-balanced plot locations, and locations of clipIDs within plots are also random. The sampling design provides an unbiased estimate of root biomass throughout the Tower airshed.

* For each `sampleID` collected in the field, test subsampling efficiency at various levels of sorting by creating subsamples with the following percentages of the total core mass:
    - 10%
    - 15%
    - 25%
    - 50%
    
    The sum of all the subsamples = 100%; that is, the mass of the entire core sample.

* When subsampling is employed, estimated `dryMass` is calculated as follows: `dM = ssDM/X`, where:
    - *dM* = estimated root dryMass for a given level of subsampling
    - *X* = the proportion of the total core sample represented by the subsample (ranging from 0-1)
    
* Compare `dryMass` results calculated via subsampling with `dryMass` obtained with no subsampling, and use mixed effects models to analyze results.


##  Procedure

1.  Soil cores (6.65 cm diameter x 30 cm max depth) were collected in the field and frozen in plastic bags for processing at a later date. To generate subsamples, thaw frozen core sample and homogenize soil/root mixture in a plastic bag or tub. 

1.  Weigh the whole core and create subsamples at desired percentages (above) based on the fresh mass of the total core (roots + soil).

1.  Sort subsamples separately into root sizeCategory (0-1, 1-2, 2-10 mm) following the normal procedure in the BBC protocol, and dry and weigh sorted subsamples separately for each sizeCategory.

1.  Record the dryMass of each sorted subsample by sizeCategory for a given core sample, then recombine all subsamples within a sizeCategory and weigh to obtain the "whole core" dryMass.
    a.  Enter the "whole core" dryMass for each sizeCategory into Fulcrum for ingest as normal.
    
    
## Data Summary

``` {r load-packages, include = FALSE}

library(dplyr)
library(ggplot2)
library(glue)
library(lmerTest)
library(openxlsx)

```

``` {r data-preparation, include = FALSE}

### Read in data collected by D18/19
if (file.exists("/Users/cmeier")) {
  inputPath <- "~/Documents/gitRepositories/neon-plant-sampling/bbcProtocolDev/subsampleSortExperiments"
}

df <- openxlsx::read.xlsx(xlsxFile = glue::glue(inputPath, "2023_BBC_subsampling_D1819.xlsx", .sep = "/"),
                          sheet = "R_input",
                          colNames = TRUE)



### Calculate estimated dryMass for subsamples
df <- df %>%
  dplyr::mutate(estDryMass = dplyr::case_when(treatment == "ss10" ~ dryMass/0.1,
                                              treatment == "ss15" ~ dryMass/0.15,
                                              treatment == "ss25" ~ dryMass/0.25,
                                              treatment == "ss50" ~ dryMass/0.5,
                                              TRUE ~ dryMass)) %>%
  dplyr::mutate(estDryMass = round(estDryMass, digits = 4)) %>%
  dplyr::mutate(treatment = dplyr::case_when(treatment == "wholeCore" ~ "compCore",
                                             TRUE ~ treatment)) %>%
  dplyr::mutate(sizeCategory = stringr::str_replace(sizeCategory,
                                                    pattern = "_",
                                                    replacement = "-"))



### Generate summary data frame for ggplot2
summaryDF <- df %>%
  dplyr::group_by(siteID,
                  treatment, 
                  sizeCategory) %>%
  dplyr::summarise(N = n(),
                   mean = mean(estDryMass),
                   sd = sd(estDryMass),
                   se = sd/sqrt(N),
                   .groups = "drop")


```

``` {r results-plot}

### Create ggplot from summaryDF and display

ggplot2::ggplot(data = summaryDF,
                mapping = aes(x = treatment,
                              y = mean,
                              fill = sizeCategory)) +
  ggplot2::geom_bar(position = "dodge",
                    stat = "identity") +
  ggplot2::geom_errorbar(mapping = aes(ymin = mean - se,
                                       ymax = mean + se),
                         width = 0.2,
                         position = position_dodge(0.9)) +
  ggplot2::labs(x = "Subsampling treatment",
                y = "Dry mass (g)",
                title = "Alaska BBC subsampling") +
  ggplot2::facet_wrap(~siteID,
                      scales = "free")


```


## Results: Mixed-effects model analysis

Approach: For each site, construct a model with `sizeCategory` as the sole fixed effect where the mean is allowed to vary randomly by `sampleID`. Next, construct a model with an additional `treatment` fixed effect, with the mean similarly allowed to vary randomly by `sampleID`. Theoretically, if subsampling has no effect on estimated dryMass the model with `treatment` will not be significantly better according to AIC or p-value than the simpler model that only has `sizeCategory`. 


### DEJU results

``` {r DEJU-mixed-effects, echo = TRUE}

#   DEJU simplest mixed-effects model
m1DEJU <- lmerTest::lmer(formula = estDryMass ~ sizeCategory + (1|sampleID),
                         data = df %>% dplyr::filter(siteID == "DEJU"),
                         REML = FALSE)

#   DEJU mixed-effects model with subsampling 'treatment'
m2DEJU <- lmerTest::lmer(formula = estDryMass ~ treatment + sizeCategory + (1|sampleID),
                         data = df %>% dplyr::filter(siteID == "DEJU"),
                         REML = FALSE)

#   DEJU mixed-effects model with treatment*sizeCategory interaction
m3DEJU <- lmerTest::lmer(formula = estDryMass ~ treatment*sizeCategory + (1|sampleID),
                         data = df %>% dplyr::filter(siteID == "DEJU"),
                         REML = FALSE)

#   Model comparison
anova(m1DEJU, m2DEJU, m3DEJU)
anova(m2DEJU, m3DEJU)

```

-   The addition of `treatment` to the `m2DEJU` model adds significant explanatory value compared to `sizeCategory` alone (lower AIC and p < 0.005 compared to `m1DEJU`).
-   Takehome from `m2DEJU`: With `sizeCategory` accounted for, the estimated dryMass is not the same across all subsampling treatments.
-   Overall: The `m3DEJU` model is significantly better than the `m2DEJU` model --> effects of subsampling treatment are not the same across sizeCategory.
-   Need to determine which `treatments` are not significantly different from the dryMass for the complete core.

``` {r DEJU-best-model-summary}

#   Summary of best model output
anova(m3DEJU)
summary(m3DEJU)

```

-   Across `sizeCategory`, the `ss10` treatment generates significantly lower estimated `dryMass` compared to the complete core.
-   Visually, the 2-10 mm `sizeCategory` for the 15% subsample is much more variable and has higher average mass than the 2-10 mm roots in the complete core. However, across all sizeCategories the 15% subsample is not significantly different from the complete core.
-   Conservatively, subsampling at 25% is supported by the data and appears most similar to the complete core.


## TOOL results

``` {r TOOL-mixed-effects, echo = TRUE}

#   TOOL simplest mixed-effects model
m1TOOL <- lmerTest::lmer(formula = estDryMass ~ sizeCategory + (1|sampleID),
                         data = df %>% dplyr::filter(siteID == "TOOL"),
                         REML = FALSE)

#   TOOL mixed-effects model with subsampling 'treatment'
m2TOOL <- lmerTest::lmer(formula = estDryMass ~ treatment + sizeCategory + (1|sampleID),
                         data = df %>% dplyr::filter(siteID == "TOOL"),
                         REML = FALSE)

#   TOOL mixed-effects model with treatment*sizeCategory interaction
m3TOOL <- lmerTest::lmer(formula = estDryMass ~ treatment*sizeCategory + (1|sampleID),
                         data = df %>% dplyr::filter(siteID == "TOOL"),
                         REML = FALSE)

m4TOOL <- lmerTest::lmer(formula = estDryMass ~ treatment + sizeCategory -1 + (1|sampleID),
                         data = df %>% dplyr::filter(siteID == "TOOL"),
                         REML = FALSE)

#   Model comparison
anova(m1TOOL, m2TOOL, m3TOOL)
anova(m2TOOL, m3TOOL)

```

-   The addition of `treatment` to the `m2TOOL` model adds significant explanatory value compared to `sizeCategory` alone (lower AIC and p < 0.05 compared to `m1TOOL`).
-   Takehome from `m2TOOL`: With `sizeCategory` accounted for, the estimated dryMass is not the same across all subsampling treatments.
-   The interaction term is not significant in the `m3TOOL` model --> subsampling has similar effects across all sizeCategories.
-   Need to determine which subsampling treatments are not different from the complete core dryMass.

``` {r TOOL-best-model-summary}

#   Summary of best model output
anova(m2TOOL)
summary(m2TOOL)

```

-   The 10% subsample treatment has on average significantly lower dryMass compared to the complete core.
-   Similar to DEJU, the 2-10 mm `sizeCategory` for the 15% subsample is much more variable and has higher average mass than the 2-10 mm roots in the complete core. Across all sizeCategories, the 15% subsample is not significantly different than the complete core.
-   Conservatively, subsampling at 25% is supported by the data and appears most similar to the complete core.




