---
title: "CDW Tally Analysis: D05 STEI"
author: "Courtney Meier, Cody Flagg"
date: "23rd August 2022"
output: 
  pdf_document: 
    latex_engine: xelatex
---

# Status

**2022-08-23 update**:

- Re-ran tally rank abundance analysis following updates to `taxonID` field.
- Results: 

**2022-08-18 update**:

- D05 FE Alex Graeff identified taxonID inconsistencies between 2016 and 2021 tallies for STEI_021 (Distributed plot). Prior IDs from 2016 were used to update 2021 records for this plot for several logs. Additionally, suspect `2PLANT-H` IDs for three logs were investigated and updated.
- Re-ran tally rank abundance analysis following data corrections to determine whether CDW DST list is affected.
- Results: Data updates were made to `acceptedTaxonID` rather than `taxonID`, so no change to DST rank list.


**2022-03-14 update**:

  - Re-running analysis in 2022 after discovery that original STEI list made in 2017 included several blank taxonIDs.
  - D05 FE Alex Graeff noted that tally effort in 2019 and 2021 employed different method to identify unknown softwoods; most are Pinaceae or _Thuja occidentalis_, so many fewer '2PLANT-S' in two most recent datasets.
  - Note: Many '2PLANT' still present in 2019 for Tower plot data --> data collection/training problem in D05.

```{r setup, results='hold', warning=FALSE, message=FALSE}
## Load libraries
library(plyr)
library(dplyr)
library(ggplot2)
library(httr)
library(glue)
library(neonUtilities)

##  Define paths and other inputs
domain <- "D05"
site <- "STEI"

# Define required paths and functions
if (file.exists("~/Documents/workDocuments")){
  outpath <- paste("~/Documents/workDocuments/gitRepositories/neon-plant-sampling/cdw_tallyAnalysis/", domain, sep="")
  source("~/Documents/workDocuments/gitRepositories/neon-plant-sampling/plant_tools/get_fulcrum_pageLim.R")
}

if (file.exists("/Users/cmeier")){
  outpath <- paste("~/Documents/gitRepositories/neon-plant-sampling/cdw_tallyAnalysis/", domain, sep="")
  source("~/Documents/gitRepositories/neon-plant-sampling/plant_tools/get_fulcrum_pageLim.R")
}



```

```{r downloadFulcrumData, results='hold', warning=FALSE, message=FALSE}
## Import data from Fulcrum
# Define CDW Fulcrum query for site
cdwQuery <- URLencode(
  glue::glue(
    'SELECT domainid, siteid, plotid_parent, tallydate, eventid, volumefactor, samplingimpractical, samplingprotocolversion, lidsazimuth, targettaxapresent, logdistance, logmaxdiameter, minorlogdiameter, logid, loglength, taxonid, scientificname, decayclass FROM "4487b90c-51b3-495a-873b-16f6b6d0abf7" AS parent JOIN "4487b90c-51b3-495a-873b-16f6b6d0abf7/cdw_transect" AS child ON (parent._record_id = child._parent_id) WHERE siteid =',
    "'{site}' ORDER BY _child_record_id",
    .sep = " "
  )
)

##  Get 2021 CDW STEI Distributed plot data from Fulcrum
fulcrumCDW <- get_fulcrum_pageLim(apiToken = Sys.getenv('FULCRUM_PAT'), sql = cdwQuery)

#   Set data types and rename variables to match Portal data
fulcrumCDW$volumefactor <- as.integer(fulcrumCDW$volumefactor)
fulcrumCDW$tallydate <- as.Date(fulcrumCDW$tallydate)

fulcrumCDW <- fulcrumCDW %>%
  dplyr::rename(
    domainID = domainid,
    siteID = siteid,
    plotID = plotid_parent,
    date = tallydate,
    eventID = eventid,
    volumeFactor = volumefactor,
    samplingImpractical = samplingimpractical,
    samplingProtocolVersion = samplingprotocolversion,
    lidsAzimuth = lidsazimuth,
    targetTaxaPresent = targettaxapresent,
    logDistance = logdistance,
    logMaxDiameter = logmaxdiameter,
    logMinorDiameter = minorlogdiameter,
    logID = logid,
    logLength = loglength,
    taxonID = taxonid,
    scientificName = scientificname,
    decayClass = decayclass
  )

fulcrumCDW$plotType <- "distributed"

#   Filter to records with samplingImpractical=='OK'
fulcrumCDW <- fulcrumCDW %>%
  dplyr::filter(
    samplingImpractical == "OK"
  )


# ##  Compare 2021 Portal STEI Distributed data with Fulcrum data
# #   Reported edits do not appear in Fulcrum data
# distPortalCDW <- neonUtilities::loadByProduct(
#   dpID = "DP1.10010.001",
#   site = "STEI",
#   startdate = "2021-01",
#   enddate = "2022-08",
#   check.size = FALSE,
#   token = Sys.getenv('NEON_PAT')
# )
# 
# distPortalCDW <- distPortalCDW$cdw_fieldtally
# #--> Odd, no updates of '2PLANT-H' to 'THOC2' appear to have been made anywhere.



##  Get older CDW Tally data from Portal for Tower plots
#   Wide date range used in case Tower plot sampling went into next year (scheduled in 2019)
portalCDW <- neonUtilities::loadByProduct(
  dpID = "DP1.10010.001",
  site = "STEI",
  startdate = "2019-01",
  enddate = "2020-12",
  check.size = FALSE,
  token = Sys.getenv('NEON_PAT')
)

portalCDW <- portalCDW$cdw_fieldtally

portalCDW <- portalCDW %>%
  dplyr::select(
    domainID,
    siteID,
    plotID,
    plotType,
    date,
    eventID,
    volumeFactor,
    samplingImpractical,
    samplingProtocolVersion,
    lidsAzimuth,
    targetTaxaPresent,
    logDistance,
    logMaxDiameter,
    logMinorDiameter,
    logID,
    logLength,
    taxonID,
    scientificName,
    decayClass
  )
portalCDW$volumeFactor <- as.integer(portalCDW$volumeFactor)


##  Combine Fulcrum and Portal datasets
#   Remove taxonID == "2PLANT"; was removed from protocol in 2017, D05 has data errors
cdw <- dplyr::bind_rows(fulcrumCDW, portalCDW) %>%
  dplyr::filter(
    taxonID != "2PLANT"
  )

```


```{r data_processing, results='hold'}
## Create diameter class factor
cdw$diameterClass <- ifelse(cdw$logMaxDiameter >= 10, '>=10cm', 
                            ifelse(cdw$logMaxDiameter < 5, "2-5cm", "5-10cm"))
 
## Simplify decayclass to numeric value wrapped inside sapply(decayClassNum, "[[", 1) e.g. 'return first element of each list member'
cdw$decayClassNum <- sapply(stringr::str_split(cdw$decayClass, pattern = " "),"[[", 1)

##  Convert '2PLANT' to '2PLANT-H'
#cdw <- cdw %>% mutate(taxonID = replace(taxonID, taxonID=='2PLANT', '2PLANT-H')) #--> cannot confidently do this


## Write data file
write.csv(cdw, file = paste(outpath, paste(domain, site, "merged","cdw_rawdata_20220823.csv", sep="_"), sep = "/"))

```

```{r summary_table, results='hold', echo=FALSE}
## Filter out 'blank' transects where TTP = NO and concentrate on decayClass = 1|2 due to cutting restrictions
cdw_fil <- cdw %>% dplyr::filter(targetTaxaPresent == "Y", !is.na(taxonID))
t1 <- ddply(cdw_fil, ~siteID+taxonID+decayClassNum+diameterClass, summarize, counts = length(taxonID))

## Sort rows by siteID then by descending number of counts
t1 <- dplyr::arrange(t1, siteID, desc(counts), diameterClass, decayClassNum, taxonID)

## Calculate relative abundances
t1$totalLogs <- nrow(cdw_fil)
t1$relativeAbundance <- round((t1$counts/t1$totalLogs)*100,2)
t1$cumulativeAbundance <- round(cumsum(t1$relativeAbundance),2)

## determine raw ranking by count of DST combo within siteid
t1 <- plyr::ddply(t1, ~siteID, mutate, DSTrank = as.numeric(paste(1:length(counts))), DSTcombo = paste0(taxonID,"_", decayClassNum,"_",diameterClass))
# t2 <- ddply(cdw, ~siteid+taxonid, summarize, counts = length(taxonid))

##  Determine required sample size per DST combo, and cumulative sample size
t1$sampledDiskNum <- ifelse(t1$diameterClass == ">=10cm", 10,5)
t1$cumulativeDiskNum <- cumsum(t1$sampledDiskNum)

```

`r knitr::kable(t1)`

```{r rank_abundance_plots, results='hold', echo=FALSE}
## species abundance/total abundance
######  This was throwing an error, unsure why
ggCDW <- ggplot(data = t1, aes(x = DSTrank, y = counts, color = taxonID, shape = diameterClass)) +
  geom_point() + 
  facet_wrap(~siteID) + 
  labs(title = "Rank Abundance of decayClass x sizeCategory x taxonID (DST)")
```

### Code

```{r write_files, echo=FALSE}
##  Prepare t1 for output to protocol appendix
# Filter columns from t1 that are not required by Field Ops, and not needed in protocol appendix
t1 <- t1 %>% select(-DSTrank, -DSTcombo, -totalLogs)

# Reorder columns
t1 <- t1[c("siteID", "taxonID", "decayClassNum", "diameterClass", "relativeAbundance", "cumulativeAbundance", "sampledDiskNum", "cumulativeDiskNum")]

# Rename columns
t1 <- t1 %>% rename(decayClass=decayClassNum, sizeCategory=diameterClass)

write.csv(t1, file = paste(outpath, paste(domain, site, "tally_abundance_20220823.csv", sep = "_"), sep = "/"), row.names = F)
```

