---
title: "BBC Dilution Sampling"
author: "Courtney Meier"
date: "10/17/2016"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)

```

## Background
- Question: Can the vortexing step in the BBC Dilution SOP be omitted, and the solution be mixed directly with the plunger?
- D05 conducted a test to determine whether there is a significant difference at their site with this simplification.
- Goal: Perform simple t-test analysis to determine if there is a significant difference between methods.

```{r dataInput}
if (file.exists(
  '/Users/cmeier/Documents/neonScienceDocs/gitRepositories')){
  wdir <- "~/Documents/neonScienceDocs/gitRepositories/neonPlantSampling/bbcProtocolDev"
}

setwd(wdir)

rootDil <- read.csv("D05_dilutionSOPmethodTesting.csv", header=T, stringsAsFactors = F)

```


##  Analysis

```{r tTest}
m1 <- t.test(dryMass~type, data = rootDil)
m1

```

*Summary*

- No `dryMass` difference between two groups.
- Slightly more mass in those subsamples that were vortexed.
    - Likely that heavier particles were not brought up into suspension as easily sans vortexing.
- Continue to use vortexing method.