---
title: "D17 DHP Photo Check"
author: "Courtney Meier"
date: "3/15/2017"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(plyr)
library(dplyr)
library(stringr)

# exifPath <- "/Volumes/neon/DSF/CI Dropbox/Domain 17/TOS/LAI_RAW_images/2016"
exifPath <- "/Volumes/neon/DSF/CI Dropbox/Domain 17/TOS/LAI_RAW_images/2017"

```


```{r getExifData, echo=FALSE, eval=FALSE}
# Read DHP Exif data for selected site using 'ExifTool'
system('exiftool -T -r -common -csv -G4 SJER03 > /Users/cmeier/Documents/neonScienceDocs/gitRepositories/neonPlantSampling/dhpPhotoCheck/dhp_exifdata_D17_2016_SJER03.csv')
system('exiftool -T -r -common -csv -G4 SJER04 > /Users/cmeier/Documents/neonScienceDocs/gitRepositories/neonPlantSampling/dhpPhotoCheck/dhp_exifdata_D17_2016_SJER04.csv')
system('exiftool -T -r -common -csv -G4 SJER07 > /Users/cmeier/Documents/neonScienceDocs/gitRepositories/neonPlantSampling/dhpPhotoCheck/dhp_exifdata_D17_2017_SJER07.csv')
system('exiftool -T -r -common -csv -G4 SJER10 > /Users/cmeier/Documents/neonScienceDocs/gitRepositories/neonPlantSampling/dhpPhotoCheck/dhp_exifdata_D17_2017_SJER10.csv')
system('exiftool -T -r -common -csv -G4 SJER00 > /Users/cmeier/Documents/neonScienceDocs/gitRepositories/neonPlantSampling/dhpPhotoCheck/dhp_exifdata_D17_2017_SJER00.csv')

```

```{r compileExifData, echo=FALSE}
# List Exif files
exifFiles <- list.files(path = ".", pattern="exifdata_D")

# Collate Exif data into data frame
newExif <- data.frame()
for (i in 1:length(exifFiles)){
  tempExif <- read.csv(exifFiles[i], header=T, stringsAsFactors = F)
  newExif <- rbind(newExif, tempExif)
}
newExif <- tbl_df(newExif)

# Make grouping variable for output summary table
sourceData <- plyr::ldply(str_split(newExif$SourceFile, "/"))
newExif$boutPlot <- paste(sourceData$V1, sourceData$V2, sep = "/")

```

##  D17 SJER DHP photo quality
### ShutterSpeed summary
Count of photos by boutPlot (boutID x plotID) and observed shutter speeds.
```{r shutterSummary, echo=FALSE}
newExif %>% count(boutPlot, ShutterSpeed) %>% kable

```

### Aperture summary
Count of photos by boutPlot (boutID x plotID) and observed apertures.
```{r apertureSummary, echo=FALSE}
newExif %>% count(boutPlot, Aperture) %>% kable

```

### ISO summary
Count of photos by boutPlot (boutID x plotID) and observed ISO values.
```{r isoSummary, echo=FALSE}
newExif %>% count(boutPlot, ISO) %>% kable

```

### Results
- **Shutter Speed**: Minimum observed shutter speed across all boutPlots was 1/30s --> good.
- **Aperture**: Minimum aperture observed across all boutPlots was `r min(newExif$Aperture)` --> good.
- **ISO**: Maximum ISO observed across all boutPlots was `r max(newExif$ISO)`, and `r nrow(newExif[newExif$ISO > 12800,])` photos had ISO >= 12,800. Technicians need to be aware when ISO is exceeding the maximum allowed value of 12,800.